### 2021-11-15 04-internet.Rmd

# Internet as predictor

```{r packages}
#| results: 'hide'
# Packages

```

##### Internet: Main effects

```{r}
# Define models of outcomes
bf_anx <- bf(
  Anxiety ~ 1 + time + sex + age + i1_cw + i1_cb +
    (1 + time + sex + age + i1_cw |c| country) +
    (1 + time + sex + age + i1_cw |r| region),
  family = gaussian()
)
bf_dep <- bf(
  Depression ~ 1 + time + sex + age + i1_cw + i1_cb +
    (1 + time + sex + age + i1_cw |c| country) +
    (1 + time + sex + age + i1_cw |r| region),
  family = gaussian()
)
bf_sel <- bf(
  Selfharm ~ 1 + time + sex + age + i1_cw + i1_cb + 
    (1 + time + sex + age + i1_cw |c| country) +
    (1 + time + sex + age + i1_cw |r| region),
  family = gaussian()
)

# Fit models in background (need 12 cores)
fit_anx <- brm(
  bf_anx,
  data = gbd_10,
  file = "output/brm-gbd-time-internet-anxiety",
  chains = 4,
  cores = 4,
  iter = hmc$iter, warmup = hmc$warmup, 
  refresh = hmc$refresh,
  control = list(adapt_delta = hmc$d)
)
fit_dep <- brm(
  bf_dep,
  data = gbd_10,
  file = "output/brm-gbd-time-internet-depression",
  chains = 4,
  cores = 4,
  iter = hmc$iter, warmup = hmc$warmup, 
  refresh = hmc$refresh,
  control = list(adapt_delta = hmc$d)
)
fit_sel <- brm(
  bf_sel,
  data = gbd_10,
  file = "output/brm-gbd-time-internet-selfharm",
  chains = 4,
  cores = 4,
  iter = hmc$iter, warmup = hmc$warmup, 
  refresh = hmc$refresh,
  control = list(adapt_delta = hmc$d)
)
fits_brm_internet_main <- tibble(
  outcome = c("Anxiety", "Depression", "Selfharm"),
  fit = list(fit_anx, fit_dep, fit_sel)
)
```

Coefficients

```{r}
fits_brm_internet_main %>% 
  mutate(
    out = map(
      fit, 
      ~as_draws(.x, variable = "b_", regex = TRUE) %>% 
        summarise_draws(quantile, rhat, .args = list(probs = c(.5, .05, .95)))
    )
  ) %>% 
  select(-fit) %>% 
  unnest(out) %>% 
  filter(str_detect(variable, "_i1_cw|b$")) %>% 
  kbl(digits = 3) %>% 
  kable_minimal(full_width = FALSE)
```

###### Fitted lines

Plot of fitted lines

```{r}
draws_country_anx <- epred_draws(
  fit_anx, 
  gbd_wide %>% 
    filter(country %in% sort(unique(fit_anx$data$country))) %>% 
    mutate(time = 0),
  re_formula = ~(1 + time + sex + age + i1_cw |r| region) + 
    (1 + time + sex + age + i1_cw |c| country), 
  ndraws = 100,
  resp = "Anxiety"
)

draws_country_dep <- epred_draws(
  fit_dep, 
  gbd_wide %>% 
    filter(country %in% sort(unique(fit_dep$data$country))) %>% 
    mutate(time = 0),
  re_formula = ~(1 + time + sex + age + i1_cw |r| region) + 
    (1 + time + sex + age + i1_cw |c| country), 
  ndraws = 30,
  resp = "Depression"
)

draws_country_sel <- epred_draws(
  fit_sel, 
  gbd_wide %>% 
    filter(country %in% sort(unique(fit_sel$data$country))) %>% 
    mutate(time = 0),
  re_formula = ~(1 + time + sex + age + i1_cw |r| region) + 
    (1 + time + sex + age + i1_cw |c| country), 
  ndraws = 30,
  resp = "Selfharm"
)

draws_region_anx <- epred_draws(
  fit_anx, 
  distinct(fit_anx$data, region, sex, age, i1_cb) %>% 
    group_by(region) %>% 
    mutate(i1_cb = mean(i1_cb, na.rm = T)) %>% 
    crossing(i1_cw = seq(-4, 4, length = 10)) %>% 
    mutate(time = 0),
  re_formula = ~(1 + time + sex + age + i1_cw |r| region),
  ndraws = 30,
  resp = "Anxiety"
) %>% 
  mutate(outcome = "Anxiety")

draws_region_dep <- epred_draws(
  fit_dep, 
  distinct(fit_dep$data, region, sex, age, i1_cb) %>%
    group_by(region) %>% 
    mutate(i1_cb = mean(i1_cb, na.rm = T)) %>% 
    crossing(i1_cw = seq(-4, 4, length = 10)) %>% 
    mutate(time = 0),
  re_formula = ~(1 + time + sex + age + i1_cw |r| region),
  ndraws = 30,
  resp = "Depression"
) %>% 
  mutate(outcome = "Depression")

draws_region_sel <- epred_draws(
  fit_sel, 
  distinct(fit_sel$data, region, sex, age, i1_cb) %>% 
    group_by(region) %>% 
    mutate(i1_cb = mean(i1_cb, na.rm = T)) %>% 
    crossing(i1_cw = seq(-4, 4, length = 10)) %>% 
    mutate(time = 0),
  re_formula = ~(1 + time + sex + age + i1_cw |r| region),
  ndraws = 30,
  resp = "Selfharm"
) %>% 
  mutate(outcome = "Self-harm")

draws_region <- bind_rows(
  draws_region_anx, draws_region_dep, draws_region_sel
) %>% 
  group_by(outcome, region, i1_cw) %>% 
  median_qi(.epred)

bind_rows(
  draws_country_anx %>% mutate(outcome = "Anxiety"),
  draws_country_dep %>% mutate(outcome = "Depression"),
  draws_country_sel %>% mutate(outcome = "Self-harm")
) %>% 
  group_by(country) %>% 
  # mutate(i1_cw2 = i1_cw/10 + mean(internet, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot(aes(i1_cw, .epred)) +
  scale_x_continuous(
    "Previous year's internet adoption relative to country average in 2000-2020",
    breaks = pretty_breaks(3),
    labels = function(x) percent(x/10, 1)
  ) +
  scale_y_continuous(
    "Estimated rate",
    breaks = pretty_breaks()
  ) +
  geom_ribbon(
    data = draws_region,
    aes(ymin = .lower, ymax = .upper),
    fill = "dodgerblue4", size = .6, alpha = .2
  ) +
  geom_line(
    data = draws_region,
    col = "dodgerblue4", size = .6
  ) +
  stat_summary(
    fun=mean, geom = "line",
    aes(group = country), 
    size = .4, alpha = .25, col = "dodgerblue1"
  ) +
  facet_grid(outcome~region, scales = "free_y") +
  theme(panel.spacing = unit(2, "pt"))
```

###### Forest plots

figure of coefficients forest plot

```{r}
anx_draws <- spread_draws(
  fit_anx, 
  b_Anxiety_i1_cw, r_region__Anxiety[region, term] | region
) %>% 
  filter(term == "i1_cw") %>% 
  ungroup() %>% 
  select(-term) %>% 
  rename(Average = b_Anxiety_i1_cw) %>% 
  mutate(outcome = "Anxiety")
dep_draws <- spread_draws(
  fit_dep, 
  b_Depression_i1_cw, r_region__Depression[region, term] | region
) %>% 
  filter(term == "i1_cw") %>% 
  ungroup() %>% 
  select(-term) %>% 
  rename(Average = b_Depression_i1_cw) %>% 
  mutate(outcome = "Depression")
sel_draws <- spread_draws(
  fit_sel, 
  b_Selfharm_i1_cw, r_region__Selfharm[region, term] | region
) %>% 
  filter(term == "i1_cw") %>% 
  ungroup() %>% 
  select(-term) %>% 
  rename(Average = b_Selfharm_i1_cw) %>% 
  mutate(outcome = "Selfharm")

results_text <- bind_rows(anx_draws, dep_draws, sel_draws) %>% 
  group_by(outcome) %>% 
  summarise(parameters(cur_data()[-c(1:3)])) %>% 
  mutate(region = str_replace(Parameter, "\\.", " ")) %>% 
  mutate(is_average = region == "Average") %>% 
  mutate(region = if_else(region=="Average", "**Average**", region)) %>% 
  mutate(across(Median:CI_high, ~format(round(., 2), nsmall = 2)))

p_internet_main_coefs <- bind_rows(anx_draws, dep_draws, sel_draws) %>% 
  mutate(across(Africa:Southern.Asia, ~.+Average)) %>% 
  pivot_longer(Average:Southern.Asia, names_to = "region") %>% 
  mutate(region = fct_relevel(factor(region), "Average")) %>% 
  mutate(outcome = factor(outcome)) %>% 
  mutate(is_average = region == "Average") %>% 
  mutate(region = str_replace(region, "\\.", " ")) %>% 
  mutate(region = if_else(region=="Average", "**Average**", region)) %>% 
  ggplot(aes(value, region, fill = is_average, alpha = is_average)) +
  scale_fill_manual(values = c("dodgerblue1", "dodgerblue4")) +
  scale_alpha_manual(values = c(.5, .9)) +
  geom_vline(xintercept = 0, lty = 2, size = .2) +
  scale_y_discrete(expand = expansion(c(0.01, .1))) +
  scale_x_continuous(
    "Estimated association to internet adoption", 
    breaks = pretty_breaks(), 
    expand = expansion(c(.1, .5))
  ) +
  stat_halfeye(
    point_interval = NULL, normalize = "panels", height = .75
  ) +
  geom_text(
    data = results_text, alpha = 1, size = 2.75, hjust = 1, vjust = -0.2,
    aes(
      x = Inf, 
      label = str_glue("{Median} [{CI_low}, {CI_high}]")
    )
  ) +
  facet_wrap("outcome", nrow = 1, scales = "free_x") +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_markdown(),
    legend.position = "none"
  )

(wrap_elements(grid::textGrob('Negative experiences\nTBD')) |
    wrap_elements(grid::textGrob('Positive experiences\nTBD')) |
    wrap_elements(grid::textGrob('Life satisfaction\nTBD'))) /
  p_internet_main_coefs
```

##### Internet: Interactions

```{r}
# Define models of outcomes
bf_anx <- bf(
  Anxiety ~ 1 + time*sex + time*age + i1_cw*sex + i1_cw*age + i1_cb +
    (1 + time*sex + time*age + i1_cw*sex + i1_cw*age | country) +
    (1 + time*sex + time*age + i1_cw*sex + i1_cw*age | region),
  family = gaussian()
)
bf_dep <- bf(
  Depression ~ 1 + time*sex + time*age + i1_cw*sex + i1_cw*age + i1_cb +
    (1 + time*sex + time*age + i1_cw*sex + i1_cw*age | country) +
    (1 + time*sex + time*age + i1_cw*sex + i1_cw*age | region),
  family = gaussian()
)
bf_sel <- bf(
  Selfharm ~ 1 + time*sex + time*age + i1_cw*sex + i1_cw*age + i1_cb + 
    (1 + time*sex + time*age + i1_cw*sex + i1_cw*age | country) +
    (1 + time*sex + time*age + i1_cw*sex + i1_cw*age | region),
  family = gaussian()
)

# Fit models
fit_anx <- brm(
  bf_anx,
  data = gbd_wide,
  file = "output/brm-gbd-time-internet-anxiety-int2",
  chains = 4,
  cores = 4,
  iter = hmc$iter, warmup = hmc$warmup, 
  refresh = hmc$refresh,
  control = list(adapt_delta = hmc$d)
)
fit_dep <- brm(
  bf_dep,
  data = gbd_wide,
  file = "output/brm-gbd-time-internet-depression-int2",
  chains = 4,
  cores = 4,
  iter = hmc$iter, warmup = hmc$warmup, 
  refresh = hmc$refresh,
  control = list(adapt_delta = hmc$d)
)
fit_sel <- brm(
  bf_sel,
  data = gbd_wide,
  file = "output/brm-gbd-time-internet-selfharm-int2",
  chains = 4,
  cores = 4,
  iter = hmc$iter, warmup = hmc$warmup, 
  refresh = hmc$refresh,
  control = list(adapt_delta = hmc$d)
)
fits_brm_internet_interactions <- tibble(
  outcome = c("Anxiety", "Depression", "Selfharm"),
  fit = list(fit_anx, fit_dep, fit_sel)
)
```

Coefficients

```{r}
fits_brm_internet_interactions %>% 
  mutate(
    out = map(
      fit, 
      ~as_draws(.x, variable = "b_", regex = TRUE) %>% 
        summarise_draws(quantile, rhat, .args = list(probs = c(.5, .05, .95)))
    )
  ) %>% 
  select(-fit) %>% 
  unnest(out) %>% 
  # filter(str_detect(variable, "_i1_cw|b$")) %>% 
  kbl(digits = 3) %>% 
  kable_minimal(full_width = FALSE)
```

Model comparisons

```{r}
tmp <- left_join(
  fits_brm_internet_main,
  fits_brm_internet_interactions %>% rename(fit_i = fit)
) %>% 
  mutate(comp = map2(fit, fit_i, waic))

```

###### Forest plots

figure of coefficients forest plot

```{r}
anx_draws <- spread_draws(
  fit_anx, 
  b_i1_cw, r_region[region, term] | region
) %>% 
  filter(term == "i1_cw") %>% 
  ungroup() %>% 
  select(-term) %>% 
  rename(Average = b_i1_cw) %>% 
  mutate(outcome = "Anxiety")
dep_draws <- spread_draws(
  fit_dep, 
  b_i1_cw, r_region[region, term] | region
) %>% 
  filter(term == "i1_cw") %>% 
  ungroup() %>% 
  select(-term) %>% 
  rename(Average = b_i1_cw) %>% 
  mutate(outcome = "Depression")
sel_draws <- spread_draws(
  fit_sel, 
  b_i1_cw, r_region[region, term] | region
) %>% 
  filter(term == "i1_cw") %>% 
  ungroup() %>% 
  select(-term) %>% 
  rename(Average = b_i1_cw) %>% 
  mutate(outcome = "Selfharm")

results_text <- bind_rows(anx_draws, dep_draws, sel_draws) %>% 
  group_by(outcome) %>% 
  summarise(parameters(cur_data()[-c(1:3)])) %>% 
  mutate(region = str_replace(Parameter, "\\.", " ")) %>% 
  mutate(is_average = region == "Average") %>% 
  mutate(region = if_else(region=="Average", "**Average**", region)) %>% 
  mutate(across(Median:CI_high, ~format(round(., 2), nsmall = 2)))

p_internet_main_coefs <- bind_rows(anx_draws, dep_draws, sel_draws) %>% 
  mutate(across(Africa:Southern.Asia, ~.+Average)) %>% 
  pivot_longer(Average:Southern.Asia, names_to = "region") %>% 
  mutate(region = fct_relevel(factor(region), "Average")) %>% 
  mutate(outcome = factor(outcome)) %>% 
  mutate(is_average = region == "Average") %>% 
  mutate(region = str_replace(region, "\\.", " ")) %>% 
  mutate(region = if_else(region=="Average", "**Average**", region)) %>% 
  ggplot(aes(value, region, fill = is_average, alpha = is_average)) +
  scale_fill_manual(values = c("dodgerblue1", "dodgerblue4")) +
  scale_alpha_manual(values = c(.4, .8)) +
  geom_vline(xintercept = 0, lty = 2, size = .2) +
  scale_shape_manual(values = c(21, 19)) +
  scale_y_discrete(expand = expansion(c(0.01, .1))) +
  scale_x_continuous(
    "Magnitude of association between internet adoption and well-being", 
    breaks = pretty_breaks(), 
    expand = expansion(c(.1, .5))
  ) +
  stat_halfeye(
    point_interval = NULL, normalize = "panels", 
    height = .75, adjust = 1.5
  ) +
  stat_pointinterval(
    alpha = 1, .width = c(.9),
    interval_size = .5,
    point_size = 1.5,
    fill = "white",
    aes(shape = after_stat(xmin > 0 | xmax < 0)),
    position = position_nudge(0, -.025)
  ) +
  geom_text(
    data = results_text, alpha = 1, size = 2.75, hjust = 1, vjust = -0.2,
    aes(
      x = Inf, 
      label = str_glue("{Median} [{CI_low}, {CI_high}]")
    )
  ) +
  facet_wrap("outcome", nrow = 1, scales = "free_x") +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_markdown(),
    legend.position = "none"
  )

(wrap_elements(grid::textGrob('Negative experiences\nTBD')) |
    wrap_elements(grid::textGrob('Positive experiences\nTBD')) |
    wrap_elements(grid::textGrob('Life satisfaction\nTBD'))) /
  p_internet_main_coefs
```

Between country?

```{r}
pred_data <- gbd_wide %>% 
  distinct(region, country, sex, age, i1_cb, i1_cw = 0, time = 0) %>% 
  filter(country %in% sort(unique(fit_anx$data$country)))

tmp <- epred_draws(
  fit_anx, pred_data, re_formula = 
    ~(1 + time*sex + time*age + i1_cw*sex + i1_cw*age | region) + 
    (1 + time*sex + time*age + i1_cw*sex + i1_cw*age | country)
)

tmp %>% 
  group_by(region, country, i1_cb) %>% 
  summarise(mean(.epred))

conditional_effects(fit_anx, "i1_cb")

gbd_wide %>% 
  group_by(country) %>% 
  summarise(i1_cb = unique(i1_cb), anx = mean(Anxiety)) %>% 
  ggplot(aes(i1_cb, anx)) +
  geom_point() +
  stat_smooth()

filter(country %in% sort(unique(fit_anx$data$country))) %>% 
  mutate(time = 0) %>% 
  group_by(country) %>% 
  mutate(xmin = min(i1_cw, na.rm = TRUE), xmax = max(i1_cw, na.rm = TRUE)) %>% 
  filter(i1_cw %in% c(xmin, xmax)) %>% 
  ungroup()
```

Sex and Age

```{r}
# Get some stuff and transform :)
get_draws <- function(fit) {
  draws_region <- spread_draws(
    fit, 
    b_i1_cw, `b_age1:i1_cw`, `b_sex1:i1_cw`,
    r_region[region, term] | term
  ) %>% 
    mutate(
      `15 to 19` = b_i1_cw + i1_cw + (`age1:i1_cw` + `b_age1:i1_cw`)*-0.5,
      `20 to 24` = b_i1_cw + i1_cw + (`age1:i1_cw` + `b_age1:i1_cw`)*0.5,
      Female = b_i1_cw + i1_cw + (`sex1:i1_cw` + `b_sex1:i1_cw`)*-0.5,
      Male = b_i1_cw + i1_cw + (`sex1:i1_cw` + `b_sex1:i1_cw`)*0.5,
      Age = `age1:i1_cw` + `b_age1:i1_cw`,
      Sex = `sex1:i1_cw` + `b_sex1:i1_cw`
    ) %>% 
    select(region, `15 to 19`:Sex)

  draws_avg <- spread_draws(
    fit, 
    b_i1_cw, `b_age1:i1_cw`, `b_sex1:i1_cw`
  ) %>% 
    mutate(
      `15 to 19` = b_i1_cw + `b_age1:i1_cw`*-0.5,
      `20 to 24` = b_i1_cw + `b_age1:i1_cw`*0.5,
      Female = b_i1_cw + `b_sex1:i1_cw`*-0.5,
      Male = b_i1_cw + `b_sex1:i1_cw`*0.5,
      Age = `b_age1:i1_cw`,
      Sex = `b_sex1:i1_cw`
    ) %>% 
    mutate(region = "**Average**") %>% 
    select(region, `15 to 19`:Sex)

  draws <- bind_rows(
    "regions" = draws_region,
    "average" = draws_avg, 
    .id = "Level"
  ) %>% 
    mutate(region = str_replace(region, "\\.", "\n")) %>% 
    ungroup()
  draws
}

draws <- bind_rows(
  "Anxiety" = get_draws(fit_anx),
  "Depression" = get_draws(fit_dep),
  "Selfharm" = get_draws(fit_sel),
  .id = "outcome"
)

draws_sum <- draws %>% 
  select(outcome, region, Age, Sex) %>% 
  group_by(outcome, region) %>% 
  summarise(
    describe_posterior(
      cur_data(), centrality = "mean", test = "pd"
    )
  ) %>% 
  rename(panel = Parameter) %>% 
  mutate(across(where(is.numeric), ~round(., 2)))

draws %>% 
  select(outcome, Level, region, `15 to 19`:Male) %>% 
  pivot_longer(-c(region, Level, outcome), names_to = "Group") %>% 
  mutate(panel = if_else(str_detect(Group, "ale"), "Sex", "Age")) %>% 
  mutate(
    side = if_else(Group %in% c("15 to 19", "Female"), "top", "bottom")
  ) %>% 
  mutate(region = str_replace(region, "\\.", " ")) %>% 
  mutate(region = fct_relevel(region, "**Average**")) %>% 
  ggplot(aes(value, region, col = Group, fill = Group)) +
  scale_alpha_manual(values = c(.6, .3)) +
  geom_vline(xintercept = 0, lty = 2, size = .2) +
  scale_shape_manual(
    values = c(21, 19), guide = "none"
    ) +
  scale_y_discrete(expand = expansion(c(0.1, 0.1))) +
  scale_x_continuous(
    "Magnitude of association between internet adoption and well-being", 
    breaks = pretty_breaks(), 
    expand = expansion(c(0, .05))
  ) +
  stat_halfeye(
    point_interval = NULL, normalize = "panels", 
    height = .75, adjust = 1.5,
    aes(side = side, group = side, alpha = Level), 
    show.legend = FALSE
  ) +
  stat_pointinterval(
    alpha = 1, .width = c(.9),
    interval_size = .33,
    point_size = 1.25,
    fill = "white",
    aes(shape = after_stat(xmin > 0 | xmax < 0), group = side),
    position = position_dodgev(.25), key_glyph = draw_key_point
  ) +
  geom_richtext(
    data = draws_sum %>% 
      mutate(Group = NA) %>% 
      filter(outcome %in% c("Anxiety", "Depression", "Selfharm")),
    aes(
      x = Inf, 
      label = str_glue("*d* = {Mean} (p<sup>d</sup> = {percent(pd, 1)})")
    ),
    size = 2.5, hjust = 1,
    family = Font,
    fill = NA, label.color = NA,
    show.legend = FALSE
  ) +
  facet_grid2(panel ~ outcome, scales = "free_x", axes = "x") +
  guides(
    alpha = "none"
  ) +
  theme(
    legend.position = "right",
    axis.title.y = element_blank(),
    axis.text.y = element_markdown()
  )
```

###### Fitted lines

Plot of fitted lines

```{r}
# Prediction dataframe: Min and max vals are enough for line
pred_data <- gbd_wide %>% 
  filter(country %in% sort(unique(fit_anx$data$country))) %>% 
  mutate(time = 0) %>% 
  group_by(country) %>% 
  mutate(xmin = min(i1_cw, na.rm = TRUE), xmax = max(i1_cw, na.rm = TRUE)) %>% 
  filter(i1_cw %in% c(xmin, xmax)) %>% 
  ungroup()
draws_country_anx <- epred_draws(
  fit_anx, 
  pred_data,
  re_formula = 
    ~(1 + time*sex + time*age + i1_cw*sex + i1_cw*age | region) + 
    (1 + time*sex + time*age + i1_cw*sex + i1_cw*age | country), 
  ndraws = 100,
  resp = "Anxiety"
)

draws_country_dep <- epred_draws(
  fit_dep, 
  pred_data,
  re_formula = 
    ~(1 + time*sex + time*age + i1_cw*sex + i1_cw*age | region) + 
    (1 + time*sex + time*age + i1_cw*sex + i1_cw*age | country), 
  ndraws = 100,
  resp = "Depression"
)

draws_country_sel <- epred_draws(
  fit_sel, 
  pred_data,
  re_formula = 
    ~(1 + time*sex + time*age + i1_cw*sex + i1_cw*age | region) + 
    (1 + time*sex + time*age + i1_cw*sex + i1_cw*age | country), 
  ndraws = 100,
  resp = "Selfharm"
)

pred_data_region <- pred_data %>% 
  distinct(region, sex, age, i1_cb) %>% 
  group_by(region) %>% 
  mutate(i1_cb = mean(i1_cb, na.rm = T)) %>% 
  crossing(i1_cw = seq(-4, 4, length = 10)) %>% 
  mutate(time = 0)

draws_region_anx <- epred_draws(
  fit_anx, 
  pred_data_region,
  re_formula = 
    ~(1 + time*sex + time*age + i1_cw*sex + i1_cw*age | region),
  ndraws = 100,
  resp = "Anxiety"
) %>% 
  mutate(outcome = "Anxiety")

draws_region_dep <- epred_draws(
  fit_dep, 
  pred_data_region,
  re_formula = 
    ~(1 + time*sex + time*age + i1_cw*sex + i1_cw*age | region),
  ndraws = 100,
  resp = "Depression"
) %>% 
  mutate(outcome = "Depression")

draws_region_sel <- epred_draws(
  fit_sel, 
  pred_data_region,
  re_formula = 
    ~(1 + time*sex + time*age + i1_cw*sex + i1_cw*age | region),
  ndraws = 100,
  resp = "Selfharm"
) %>% 
  mutate(outcome = "Self-harm")

means_country <- bind_rows(
  draws_country_anx %>% mutate(outcome = "Anxiety"),
  draws_country_dep %>% mutate(outcome = "Depression"),
  draws_country_sel %>% mutate(outcome = "Self-harm")
) %>% 
  group_by(outcome, region, country, i1_cw) %>% 
  summarise(.epred = mean(.epred))

means_region <- bind_rows(
  draws_region_anx, draws_region_dep, draws_region_sel
) %>% 
  group_by(outcome, region, i1_cw) %>% 
  mean_qi(.epred, .width = .9)

means_region <- pseudo_gwp %>% 
  distinct(region, outcome=cause) %>% 
  bind_rows(means_region) %>% 
  mutate(outcome = fct_inorder(factor(outcome)))

pseudo_gwp %>% 
  distinct(region, outcome=cause) %>% 
  bind_rows(means_country) %>% 
  mutate(outcome = fct_inorder(factor(outcome))) %>% 
  ggplot(aes(i1_cw, .epred)) +
  scale_x_continuous(
    "Previous year's internet adoption relative to country average in 2000-2020",
    breaks = c(-3.5, 0, 3.5),
    labels = function(x) percent(x/10, 1)
  ) +
  scale_y_continuous(
    "Estimated rate or scale score",
    breaks = pretty_breaks()
  ) +
  geom_blank() +
  geom_ribbon(
    data = means_region,
    aes(ymin = .lower, ymax = .upper),
    alpha = .15, fill = "dodgerblue4"
  ) +
  geom_line(
    data = means_region,
    col = "dodgerblue4", size = .6
  ) +
  geom_line(
    aes(group = country), 
    size = .4, alpha = .35, col = "dodgerblue1"
  ) +
  # geom_text(
  #   data = bind_rows(anx_draws, dep_draws, sel_draws) %>% 
  #     group_by(outcome) %>% 
  #     summarise(parameters(cur_data()[-c(1:3)])) %>% 
  #     mutate(region = str_replace(Parameter, "\\.", "\n")) %>% 
  #     mutate(outcome = if_else(outcome=="Selfharm", "Self-harm", outcome)) %>% 
  #     filter(region != "Average") %>% 
  #     mutate(across(where(is.numeric), ~round(., 2))),
  #   aes(x = -Inf, y = Inf, label = str_glue("{Median}\n[{CI_low}, {CI_high}]")),
  #   col = "dodgerblue4", size = 2, vjust = 1.1, hjust = -0.1, family = Font
  # ) +
facet_grid(outcome~region, scales = "free_y") +
  theme(panel.spacing = unit(2, "pt"))
```