### 2022-01-10 index.qmd

# Preface {.unnumbered}

Welcome! This document is the online analysis supplement to *A global study of the association between internet adoption and adolescent mental health and well-being* (Vuorre & Przybylski, [Preprint]() in preparation). It contains all the analyses reported in the manuscript along with supplemental analyses. The data and source code repository, along with the datasets, are on [GitHub](https://github.com/digital-wellbeing/global-wbmh), and an archived permanent copy of these are on [OSF](https://osf.io/ys7m9/).

::: {.callout-note}
### How to reproduce our results

The analyses were conducted in R, and can be reproduced locally:

1. Clone the github repo
  - Terminal: `git clone https://github.com/digital-wellbeing/global-wbmh.git`  
      OR
  - RStudio: File -> New Project -> Version Control -> Git -> use the URL from above

2. Prepare the R environment
  - Terminal: `Rscript -e "renv::restore()"`  
      OR
  - RStudio: Click renv -> Restore Library in the Packages panel

3. Render the source files to a Quarto book
 - Install [Quarto](https://quarto.org/docs/getting-started/installation.html) if you don't already have it
  - Terminal: `quarto render`  
      OR
  - RStudio: Open the project (`global-wbmh.Rproj`) and click Render in the Build panel

::: {.callout-important}
The project repo includes the GBD and ITU datasets, and **a synthetic mock version of the GWP dataset** to enable reproducing all our computations.
**The models take several hours each to run**---depending on your local computing resources---and therefore the rendering process can take several days. For this reason it might be easier to run the individual `.qmd` files interactively in RStudio.
:::

If something doesn't work please [open an issue](https://github.com/digital-wellbeing/global-wbmh/issues).
:::