### 2022-02-02 timecourses.Rmd
Revise for new RE structure

# Time courses

```{r setup}
#| cache: false
# Packages
library(scales)
library(ggtext)
library(lubridate)
library(kableExtra)
library(ggstance)
library(officer)
library(bayestestR)
library(ellipse)
library(ggh4x)
library(mgcv)
library(tidybayes)
library(ggdist)
library(posterior)
library(parameters)
library(emmeans)
library(broom)
library(brms)
library(tidyverse)

# Load data sets
itu <- read_rds("data/itu.rds")
gwp <- read_rds("data/gwp.rds")
gbd <- read_rds("data/gbd.rds")

# Load models
oo <- c(
  "Life_satisfaction", "Negative_experiences", "Positive_experiences",
  "Anxiety", "Depression", "Selfharm"
)
fits <- tibble(outcome = oo) %>% 
  slice(1:3) %>% 
  mutate(outcome = factor(outcome, levels = oo)) %>% 
  mutate(
    fit1 = map(
      outcome, 
      ~read_rds(str_glue("models/brm-{.x}-1-internet.rds"))
    )
  )
```

Our first research question concerns changes in mental health and well-being over time: To what extent has adolescent mental health changed over time? And, do changes in well-being correlate with changes in internet adoption? To answer, we examine Model 1's results.

## Data figure

```{r trends-regions}
#| echo: false
#| fig-width: 12
#| fig-height: 8
#| fig-cap: "Trends in all variables across regions"
#| label: fig-trends-regions

# Stack data sets
# Assume all outcomes are proportions (0-1)
# Draw small multiples figure, outcomes x regions
bind_rows(
  gwp,
  gbd,
  itu %>% mutate(outcome = "Internet", val = internet)
) %>%
  mutate(outcome = fct_inorder(str_replace(outcome, "_", "_"))) %>%
  group_by(region, country, year, outcome) %>%
  summarise(mean = mean(val, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(year = make_date(year)) %>%
  ggplot(aes(year, mean)) +
  scale_y_continuous(
    "Value",
    labels = function(x) percent(x, .01) %>% str_remove("\\.00"),
    breaks = pretty_breaks()
  ) +
  scale_x_date(
    "Year",
    date_breaks = "5 years",
    date_labels = "'%y",
    expand = expansion(.02)
  ) +
  geom_line(
    aes(group = interaction(country, outcome)),
    alpha = .25, size = .4, color = "dodgerblue1"
  ) +
  stat_summary(
    fun = mean, geom = "line",
    size = .6, color = "dodgerblue4"
  ) +
  facet_grid(
    outcome ~ region,
    scales = "free_y",
    labeller = as_labeller(function(x) str_replace(x, "_", "\n"))
  ) +
  theme(
    aspect.ratio = 1,
    legend.position = "none",
    panel.spacing = unit(4, "pt"),
    axis.title.y = element_blank()
  )
```

```{r}
tmp <- itu %>% 
  arrange(country, year) %>% 
  group_by(country) %>% 
  mutate(i = internet - lag(internet)) %>% 
  summarise(i = mean(i, na.rm = TRUE))
internet_yoy <- percent(mean(tmp$i, na.rm = TRUE), .1)
```

Average year-on-year increase in internet adoption was `r internet_yoy`.

## Main result: Trends

```{r}
get_tab <- function(pattern, label, sort = outcome) {
  fits %>% 
    mutate(
      out = map(
        fit1,
        ~as_draws_df(
          .x, 
          variable = pattern, 
          regex = TRUE
        ) %>% 
          summarise_draws(
            mean, 
            ~quantile2(., probs = c(.025, .975)),
            pd = ~pd(as.numeric(.)), 
            rhat
            )
      )
    ) %>% 
    select(outcome, out) %>% 
    unnest(out) %>% 
    mutate(pd = percent(pd, .1)) %>% 
    arrange({{sort}}) %>% 
    kbl(digits = 2, caption = label) %>% 
    kable_custom()
}

get_tab(
  "b_val_(Intercept|year)$", 
  "Model 1: Population level effects",
  variable
)
```

This table shows the population-level (average across region, country, age) intercepts and time-courses. Negative experiences had increased. We then summarise the region- and country-level trends in a table.

```{r}
get_coeftab <- function(x, group) {
  coef(x, summary = FALSE)[[group]][,,"val_year"] %>% 
    summarise_draws(
      mean, 
      ~quantile2(., probs = c(.025, .975)),
      pd = ~pd(as.numeric(.)), 
      rhat
    ) %>% 
    mutate(pd = percent(pd, .1))
}

val_year_region <- fits %>% 
  mutate(out = map(fit1, ~get_coeftab(.x, "region"))) %>% 
  select(outcome, out)
val_year_country <- fits %>% 
  mutate(out = map(fit1, ~get_coeftab(.x, "country"))) %>% 
  select(outcome, out)

bind_rows(
  "region" = val_year_region,
  "country" = val_year_country,
  .id = "Grouping"
) %>% 
  unnest(out) %>% 
  rename(Group = variable) %>% 
  kbl(digits = 2, caption = "Country-specific timecourses") %>% 
  kable_custom() %>% 
  scroll_box("500px")
```

Before presenting the time trends, we must calculate the required quantities from the posterior distribution. Here we calculate posterior samples of region and average-level trends over time, and posterior summaries of trends at all levels.

```{r calculate-simple-effects}
p_draws <- fits %>%
  mutate(pdraws = map(fit1, pdraws)) %>%
  select(outcome, pdraws) %>%
  unnest(pdraws) %>%
  mutate(region = fct_relevel(region, "Average"))
p_sums <- fits %>%
  mutate(psums = map(fit1, psums)) %>%
  select(outcome, psums) %>%
  unnest(psums) %>%
  mutate(region = fct_relevel(region, "Average"))
p_sums_internet <- fits %>%
  mutate(psums = map(fit1, ~psums(.x, "internet_year"))) %>%
  select(outcome, psums) %>%
  unnest(psums) %>%
  mutate(region = fct_relevel(region, "Average"))
```

We then display the average ($\beta^y_1$), region- ($\delta^y_1$), and country-level ($\gamma^y_1$ in @eq-model1) time trend parameters' posterior distributions. Recall the scales of the variables when interpreting the magnitudes: Year is in decades and outcomes are percentages (0-100).

```{r forest-trends}
#| column: screen-inset-right
#| fig-width: 13
#| fig-height: 8
#| fig-cap: "Forest plot of linear trends in outcomes. Shaded areas indicate approximate posterior densities of average and region-specific trends. Blue points are country-specific posterior means. Black points and lines are 95%CIs; open points indicate estimates where zero is within the 95% most credible values."
#| label: fig-forest-trends
p_draws %>%
  mutate(is_average = region == "Average") %>%
  ggplot(aes(value, region, fill = is_average)) +
  scale_fill_manual(values = c("dodgerblue1", "dodgerblue4")) +
  scale_alpha_manual(values = c(.4, .8)) +
  geom_vline(xintercept = 0, lty = 2, size = .2) +
  scale_shape_manual(values = c(21, 19)) +
  scale_y_discrete(
    expand = expansion(c(0.01, .1)),
    labels = function(x) {
      if_else(x == "Average", "**Average**", x) %>%
        str_replace("_", " ")
    }
  ) +
  scale_x_continuous(
    "Estimated linear change per decade",
    breaks = pretty_breaks(),
    labels = function(x) percent(x / 100, .01) %>% str_remove("\\.00"),
    expand = expansion(c(.1, .1))
  ) +
  stat_halfeye(
    point_interval = NULL, normalize = "panels",
    height = .75, adjust = 1,
    aes(alpha = is_average)
  ) +
  # Add country points
  geom_point(
    data = p_sums %>%
      drop_na(country) %>%
      mutate(is_average = TRUE),
    aes(x = Estimate), size = 1, col = "dodgerblue1",
    position = position_nudge(0, -0.075), alpha = .4
  ) +
  stat_pointinterval(
    alpha = 1, .width = c(.95),
    interval_size = 1,
    point_size = 1.75,
    fill = "white",
    aes(shape = after_stat(xmin > 0 | xmax < 0)),
    position = position_nudge(0, 0)
  ) +
  geom_richtext(
    data = p_sums %>%
      filter(is.na(country)) %>%
      mutate(is_average = region == "Average") %>%
      mutate(
        across(
          where(is.numeric),
          ~ if_else(
            outcome == "Selfharm",
            percent(. / 100, .001),
            percent(. / 100, .01)
          )
        )
      ),
    family = Font,
    alpha = 1, size = 3, hjust = 1, vjust = 0.2,
    fill = NA, label.color = NA,
    aes(
      x = Inf,
      label = str_glue("{Estimate}<br>[{Q2.5}, {Q97.5}]")
    )
  ) +
  facet_wrap(
    "outcome",
    nrow = 1, scales = "free_x",
    labeller = as_labeller(~ str_replace(., "_", " "))
  ) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_markdown(),
    legend.position = "none"
  )
```

Starting from the top left panel, the average (across countries and regions) change in life satisfaction. This change is not credibly different from zero, at the 95% credibility level, and the posterior mean value of change was a 1% increase in life satisfaction per decade. The region-specific changes were likewise not credibly different from zero, apart from the Eastern Europe region that indicated a modest increase in life satisfaction.

Because we modelled regions (and countries) as draws from a multivariate normal prior distribution, the region-specific estimates are shrunk toward the overall mean and therefore less likely to overfit any individual region's data. This means that the model is less likely to report "false alarms", or estimates at the region- and country-levels that indicate false certainty in nonzero changes. Therefore, although there are a large number of change parameters for each individual outcome, we can be fairly certain that the observed change in Eastern Europe is not due to sampling error but represents a real trend in that population.

We observe credible increases in negative experiences in more regions, with notably a \~10% increase per decade in the African region. However, the average change again is not credibly different from zero. Positive experiences had increased, both on average and across 7/10 regions, with very similar changes in Anxiety. Depression and self-harm, on the other hand, had decreased across the board.

In sum these results were somewhat mixed, where some regions showing both increases (e.g. decreased depression) and decreases (increased anxiety) in well-being.

Then to help illustrate the parameters, we draw figures of the fitted lines conditional on the predictors, or so called counterfactual plots.

## Main result: Correlations

We then turned to research question 1b: To what extent were changes in well-being and internet adoption correlated across countries.

To investigate we examined the correlations between $\gamma^y_1$ and $\gamma^x_1$ ($\sigma_{\gamma^y_1,\gamma^x_1}$ in $\Sigma^{\text{c}}$). A positive correlation would indicate that countries with greater increases in internet adoption between 2000 and 2020 would tend to have greater increases in $y$ (e.g. Life satisfaction).

To answer, we display $\sigma_{\gamma^y_1,\gamma^x_1}$ from each outcome's model in the figure below. The figures are scatterplots of country-level trend coefficients, where each point's x-y coordinate indicate their trend magnitudes. The overlaid ellipses are the model's 95% predictive limits for $\gamma^y_1$ and $\gamma^x_1$.

```{r scatterplot-time-internet}
#| column: screen-inset-right
#| fig-width: 12
#| fig-height: 8
#| fig-cap: "Scatterplots of country-level internet and outcome trends."
#| label: fig-trend-scatterplot

# Correlation parameters
cor_pars <- fits %>%
  mutate(
    out = map(
      fit1,
      ~ as_draws_df(
        .x,
        variable = "cor_country__internet_year__val_year"
      ) %>%
        summarise_draws(mean, sd, ~ quantile2(., probs = c(.025, .975)))
    )
  ) %>%
  select(outcome, out) %>%
  unnest(out) %>%
  mutate(across(where(is.numeric), ~ number(., .01)))

# Display table and print in word document
cor_tab <- cor_pars %>% 
  transmute(
    outcome = str_replace(outcome, "_", " "), 
    correlation = str_glue("{mean} [{q2.5}, {q97.5}]")
  )
cor_tab %>%   
  kbl(caption = "Between-country trend correlation parameters") %>% 
  kable_custom()

read_docx() %>% 
  body_add_table(
    cor_tab,
    first_column = TRUE
  ) %>% 
  print("Table-correlations.docx")

# Ellipses
ellipses <- fits %>%
  mutate(
    ellipse = map(
      fit1,
      ~ as_draws_df(
        .x,
        c(
          "b_val_year",
          "b_internet_year",
          "sd_country__val_year",
          "sd_country__internet_year",
          "cor_country__internet_year__val_year"
        )
      ) %>%
        slice_sample(n = 100) %>%
        rename(
          cor = cor_country__internet_year__val_year,
          sd_outcome = sd_country__val_year,
          sd_i = sd_country__internet_year,
          b_outcome = b_val_year,
          b_i = b_internet_year
        ) %>%
        mutate(
          e = pmap(
            list(cor, sd_outcome, sd_i, b_outcome, b_i),
            function(cor, sd_a, sd_i, b_a, b_i) {
              ellipse(
                cor,
                scale = c(sd_i, sd_outcome),
                centre = c(b_i, b_outcome),
                level = .95
              ) %>%
                as_tibble()
            }
          )
        ) %>%
        unnest(e)
    )
  )
ellipses <- ellipses %>%
  select(outcome, ellipse) %>%
  unnest(ellipse)

# Wide tibble of parameters
left_join(
  p_sums %>%
    drop_na(country) %>%
    rename_with(~ str_glue("Outcome_{.}"), .cols = c(Estimate:Q97.5)),
  p_sums_internet %>%
    drop_na(country) %>%
    rename_with(~ str_glue("Internet_{.}"), .cols = c(Estimate:Q97.5))
) %>%
  ggplot(aes(Internet_Estimate, Outcome_Estimate)) +
  geom_vline(xintercept = 0, lty = 2, size = .2) +
  geom_hline(yintercept = 0, lty = 2, size = .2) +
  geom_path(
    data = ellipses,
    aes(x = x, y = y, group = .draw),
    alpha = .2, size = .1, col = "dodgerblue2"
  ) +
  geom_point(col = "dodgerblue1", size = 1, alpha = .5) +
  scale_y_continuous(
    "Outcome trend",
    breaks = pretty_breaks(),
    labels = function(x) percent(x / 100, .01) %>% str_remove("\\.00")
  ) +
  scale_x_continuous(
    "Internet trend",
    breaks = pretty_breaks()
  ) +
  geom_text(
    data = cor_pars,
    aes(x = -Inf, y = Inf, label = str_glue("{mean} [{q2.5}, {q97.5}]")),
    vjust = 1.3, hjust = -0.025, family = Font
  ) +
  facet_wrap(
    "outcome",
    scales = "free_y",
    labeller = as_labeller(~ str_replace(., "_", " "))
  ) +
  theme(aspect.ratio = 1)
```

## Country-specific lines

::: {.panel-tabset .column-screen-inset-right}

### Life satisfaction

```{r trends-ls}
f_country <- fits %>%
  mutate(fc = map(fit1, fitted_country)) %>%
  select(outcome, fc)

#' Draw a panel plot of countries' fitted values and data
#'
#' @param x fitted country level data from fitted_country()
#' @param y outcome
#'
#' @return a ggplot for that outcome
plot_country <- function(x, y) {
  coefs <- p_sums %>%
    drop_na(country) %>%
    filter(outcome == y) %>%
    select(country, coef = Estimate, coef.l = Q2.5, coef.u = Q97.5) %>%
    mutate(country = fct_reorder(country, coef))
  x %>%
    left_join(coefs) %>%
    mutate(country = fct_reorder(country, coef)) %>%
    ggplot(aes(make_date(year * 10 + 2010), Estimate)) +
    scale_x_date("Year", date_breaks = "5 years", date_labels = "'%y") +
    scale_y_continuous(
      str_replace(y, "_", " "),
      breaks = pretty_breaks(),
      expand = expansion(.15),
      labels = function(x) {
        percent(x / 100, .01) %>%
          str_remove("\\.00")
      }
    ) +
    # Add parameters in text
    geom_richtext(
      data = coefs %>%
        mutate(across(where(is.numeric), ~ number(., .01))),
      aes(
        x = make_date(2000), y = Inf,
        label = str_glue("{coef}<br>[{coef.l}, {coef.u}]")
      ),
      size = 2.4, vjust = 1, hjust = 0.05, color = "black",
      family = Font,
      fill = alpha("white", .3), label.color = NA,
      label.padding = unit(c(0.1, 0.1, 0.1, 0.1), "lines")
    ) +
    geom_point(aes(y = val), shape = 1, alpha = .5) +
    geom_ribbon(
      aes(ymin = Q2.5, ymax = Q97.5),
      alpha = .3, col = NA, fill = "dodgerblue1"
    ) +
    geom_line(
      size = 0.5, col = "dodgerblue1"
    ) +
    facet_wrap(
      "country"
    ) +
    theme(
      panel.spacing = unit(2, "pt")
    )
}

# Create the plot for each outcome
f_country <- f_country %>%
  mutate(
    plot = map2(
      fc, outcome,
      ~ plot_country(.x, .y)
    )
  )
#| echo: false
#| fig-width: 12
#| fig-height: 10
#| fig-cap: "Life satisfaction trends, ordered by association magnitude"
#| label: fig-trends-ls
f_country$plot[[1]]
```

### Negative experiences

```{r trends-ne}
#| echo: false
#| fig-width: 12
#| fig-height: 10
#| fig-cap: "Negative experiences trends, ordered by association magnitude"
#| label: fig-trends-ne
f_country$plot[[2]]
```

### Positive experiences

```{r trends-pe}
#| echo: false
#| fig-width: 12
#| fig-height: 10
#| fig-cap: "Positive experiences trends, ordered by association magnitude"
#| label: fig-trends-pe
f_country$plot[[3]]
```

### Anxiety

```{r trends-anx}
#| echo: false
#| fig-width: 12
#| fig-height: 10
#| fig-cap: "Anxiety trends, ordered by association magnitude"
#| label: fig-trends-anx
f_country$plot[[4]]
```

### Depression

```{r trends-dep}
#| echo: false
#| fig-width: 12
#| fig-height: 10
#| fig-cap: "Depression trends, ordered by association magnitude"
#| label: fig-trends-dep
f_country$plot[[5]]
```

### Self-harm

```{r trends-sel}
#| echo: false
#| fig-width: 12
#| fig-height: 10
#| fig-cap: "Self-harm trends, ordered by association magnitude"
#| label: fig-trends-sel
f_country$plot[[6]]
```
:::

## Other results

### Shrinkage

It's important to note that at the level of countries and regions, the model parameters refer to country^\*^ and region^\*^, that is as they are in the sample of other countries and regions. This modelling strategy is important because it allows describing the parameters at each level while adjusting the values toward the group means (preventing "false alarms"). However it is then important to examine those values against their "fixed effect" counterparts, which we do here.

```{r shrinkage}
#| column: screen-inset-right
#| fig-width: 12
#| fig-height: 8
#| fig-cap: "Scatterplots of multilevel and independent model coefficients"
#| label: fig-trend-shrinkage

lm_coefs <- fits %>%
  mutate(
    data = map(fit1, ~ as_tibble(.x$data))
  ) %>%
  mutate(
    fit_val = map(
      data,
      ~ tidy(
        lm(val ~ 0 + country + year:country, data = .x)
      )
    ),
    fit_internet = map(
      data,
      ~ tidy(
        gam(
          internet ~ 0 + country + year:country,
          family = betar, data = .x %>%
            filter(itu)
        ),
        parametric = TRUE
      )
    )
  )
lm_coefs <- lm_coefs %>%
  select(outcome, fit_val, fit_internet) %>%
  pivot_longer(c(fit_val, fit_internet)) %>%
  mutate(name = str_remove(name, "fit_")) %>%
  unnest(value) %>%
  select(outcome, name, term, estimate, std.error) %>%
  filter(str_detect(term, ":")) %>%
  mutate(term = str_remove(term, "country")) %>%
  mutate(term = str_remove(term, ":year")) %>%
  pivot_wider(names_from = name, values_from = c(estimate, std.error)) %>%
  rename(country = term)

# Plot
left_join(
  p_sums %>%
    drop_na(country) %>%
    rename_with(~ str_glue("Outcome_{.}"), .cols = c(Estimate:Q97.5)),
  p_sums_internet %>%
    drop_na(country) %>%
    rename_with(~ str_glue("Internet_{.}"), .cols = c(Estimate:Q97.5))
) %>%
  left_join(lm_coefs) %>%
  mutate(region = str_replace(region, "_", " ")) %>%
  ggplot(aes(Internet_Estimate, Outcome_Estimate, col = region)) +
  geom_vline(xintercept = 0, lty = 2, size = .2) +
  geom_hline(yintercept = 0, lty = 2, size = .2) +
  geom_segment(
    aes(
      x = estimate_internet, xend = Internet_Estimate,
      y = estimate_val, yend = Outcome_Estimate
    ),
    arrow = arrow(type = "closed", length = unit(3, "pt")),
    size = .25, show.legend = FALSE
  ) +
  geom_point(size = 1.1, alpha = .4) +
  scale_y_continuous(
    "Outcome trend",
    breaks = pretty_breaks(),
    labels = function(x) percent(x / 100, .01) %>% str_remove("\\.00")
  ) +
  scale_x_continuous(
    "Internet trend",
    breaks = pretty_breaks()
  ) +
  scale_color_discrete("Region") +
  facet_wrap(
    "outcome",
    scales = "free_y",
    labeller = as_labeller(function(x) str_replace(x, "_", " "))
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 2, alpha = 1))
  ) +
  theme(
    aspect.ratio = 1,
    legend.position = "bottom"
  )
```

This figure indicates, for each country's internet trend and outcome trend parameters, the degree of shrinkage. That is, the start of each arrow indicates the magnitudes from a simple linear regression for that country. The end of the arrow indicates the magnitudes from the multilevel regression model. We can then see that especially for countries further away from the overall mean effects, there was considerable shrinkage. It is important to keep this in mind when interpreting the country-level parameters.