### 2022-03-24 02-correlations.Rmd

# Between-country correlations

```{r}
# Between-country correlations

library(scales)
library(memoise)
library(tidyverse)
library(kableExtra)
library(ggtext)
library(cachem)
library(bayestestR)
library(multidplyr)
library(broom)
library(brms)
library(posterior)
library(ggstance)
library(ellipse)
library(ggh4x)
library(patchwork)

cd <- cache_disk("memoise_cache", max_size = Inf)
cluster <- new_cluster(3)

dat <- readRDS("data/data-all.rds")
itu <- readRDS("data/itu.rds")

oo <- c(
  "Life_satisfaction", "Negative_experiences", "Positive_experiences",
  "Anxiety", "Depression", "Selfharm"
)
fits <- tibble(outcome = oo) %>%
  mutate(outcome = factor(outcome, levels = oo))
```

## OLS

```{r}
# Parameters for delta scatterplots
delta_ols_i <- itu %>%
  mutate(year = year - 2010) %>%
  group_by(region, country) %>%
  group_modify(~tidy(lm(internet ~ year, data = .))) %>%
  filter(term == "year") %>%
  select(country, estimate_internet = estimate)
delta_ols_m <- itu %>%
  drop_na(mobile) %>%
  mutate(year = year - 2010) %>%
  group_by(region, country) %>%
  group_modify(~tidy(lm(mobile ~ year, data = .))) %>%
  filter(term == "year") %>%
  select(country, estimate_mobile = estimate)
delta_ols_v <- dat %>%
  drop_na(val) %>%
  mutate(year = year - 2010) %>%
  group_by(region, outcome, country, year) %>%
  summarise(val = mean(val, na.rm = TRUE)) %>%
  group_modify(~tidy(lm(val ~ year, data = .))) %>%
  filter(term == "year") %>%
  select(country, estimate_val = estimate)
delta_ols <- left_join(
  delta_ols_v,
  delta_ols_i
) %>%
  left_join(delta_ols_m) %>%
  ungroup()

p_delta_ols_internet <- delta_ols %>%
  ggplot(aes(estimate_internet, estimate_val, col = region)) +
  scale_color_discreterainbow(
    name = "Region",
    labels = ~str_replace(., "_", " ")
  ) +
  scale_y_continuous(
    labels = ~percent(./100, .1)
  ) +
  scale_x_continuous(
    labels = ~percent(., .1)
  ) +
  geom_point(
    shape = 1, size = 1.0, alpha = .75
  ) +
  facet_wrap(
    "outcome",
    scales = "free_y",
    nrow = 1,
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  theme(
    aspect.ratio = 1
  ) +
  guides(
    color = guide_legend(
      override.aes = list(alpha = 1, shape = 16, size = 1.5)
    )
  ) +
  labs(
    y = expression(Delta~"Outcome"),
    x = expression(Delta~"Internet users per capita")
  )

p_delta_ols_mobile <- p_delta_ols_internet +
  aes(x = estimate_mobile) +
  labs(
    x = expression(Delta~"Mobile users per capita")
  )

# Data for average levels scatterplots
dat_avg <- dat %>%
  group_by(outcome, region, country) %>%
  summarise(
    internet = mean(internet, na.rm = TRUE),
    mobile = mean(mobile, na.rm = TRUE),
    val = mean(val, na.rm = TRUE)
  ) %>%
  ungroup()

p_means_ols_internet <- dat_avg %>%
  ggplot(aes(y = val, col = region)) +
  scale_color_discreterainbow(
    name = "Region",
    labels = ~str_replace(., "_", " ")
  ) +
  scale_y_continuous(
    "Mean outcome",
    labels = ~percent(./100, .1),
    breaks = extended_breaks()
  ) +
  geom_point(
    shape = 1, size = 1.0, alpha = .75
  ) +
  facet_wrap(
    ~outcome,
    scales = "free", nrow = 1,
    labeller = as_labeller(~str_replace(., "_", " ") %>% str_to_title(.))
  ) +
  guides(
    color = guide_legend(
      override.aes = list(alpha = 1, shape = 16, size = 1.5)
    )
  ) +
  theme(
    aspect.ratio = 1
  )

p_means_ols_internet <- p_means_ols_internet +
  aes(x = internet) +
  scale_x_continuous(
    "Mean internet users per capita",
    breaks = seq(0, 1, by = .25),
    labels = ~percent(.),
    limits = 0:1
  )
p_means_ols_mobile <- p_means_ols_internet +
  aes(x = mobile) +
  scale_x_continuous(
    "Mean mobile internet subscriptions per capita",
    breaks = seq(0, 1.5, by = .5),
    labels = ~percent(.),
    limits = c(0, 1.5)
  )

(p_means_ols_internet | 
    p_delta_ols_internet | 
    p_means_ols_mobile | 
    p_delta_ols_mobile
) +
  plot_layout(nrow = 4, guides = "collect")

agg_png(
  "figures/Fig-country-correlations-OLS.png",
  width = W, height = W*0.8, units = "in", res = 400
)
last_plot()
dev.off()

# Coefficient plot
means_ols_cors <- dat_avg %>%
  group_by(outcome) %>%
  group_modify(
    ~tidy(lm(scale(val) ~ scale(internet), data = .), conf.int = TRUE)
  ) %>%
  filter(term != "(Intercept)")
delta_ols_cors <- delta_ols %>%
  group_by(outcome) %>%
  group_modify(
    ~tidy(lm(scale(estimate_val) ~ scale(estimate_internet), data = .), conf.int = TRUE)
  ) %>%
  filter(term != "(Intercept)")
cor_pars_ols <- bind_rows(
  "Means" = means_ols_cors,
  "Deltas" = delta_ols_cors,
  .id = "type"
)
cor_pars_ols %>%
  mutate(type = fct_rev(type)) %>%
  mutate(outcome = fct_rev(factor(outcome, levels = oo))) %>%
  ggplot(aes(estimate, outcome)) +
  geom_vline(xintercept = 0, lty = 2, size = .25) +
  geom_pointrangeh(aes(xmin = conf.low, xmax = conf.high)) +
  facet_wrap("type") +
  labs(x = "Correlation") +
  theme(axis.title.y = element_blank())
```

## Model based

### Correlation parameters

```{r}
get_cor_pars <- function(outcome, predictor, model) {
  x <- readRDS(str_glue("models/brm-{outcome}-{model}.rds"))
  as_draws_df(
    x,
    variable = c(
      str_glue("b_{predictor}_Intercept"),
      "b_val_Intercept",
      str_glue("b_{predictor}_year"),
      "b_val_year",
      str_glue("sd_country__{predictor}_Intercept"),
      "sd_country__val_Intercept",
      str_glue("sd_country__{predictor}_year"),
      "sd_country__val_year",
      str_glue("cor_country__{predictor}_year__val_year"),
      str_glue("cor_country__{predictor}_Intercept__val_Intercept")
    )
  )
}
get_cor_pars <- memoise(get_cor_pars, cache = cd)

cor_pars_brm_internet <- fits %>%
  mutate(out = map(outcome, ~get_cor_pars(.x, "internet", "1")))
cor_pars_brm_mobile <- fits %>%
  mutate(out = map(outcome, ~get_cor_pars(.x, "mobile", "4")))

cor_pars_brm <- bind_rows(
  "internet" = cor_pars_brm_internet,
  "mobile" = cor_pars_brm_mobile,
  .id = "type"
) %>% 
  mutate(
    out = map(
      out,
      ~summarize_draws(., mean, ~quantile2(., probs = c(.025, .975))))
  ) %>%
  unnest(out) %>% 
  filter(str_detect(variable, "cor_country")) %>% 
  mutate(variable = if_else(str_detect(variable, "Intercept"), "Intercepts", "Slopes")) %>% 
  mutate(outcome = factor(outcome, levels = oo))
cor_pars_brm %>%   
  ggplot(aes(mean, outcome, col = type)) +
  geom_vline(xintercept = 0, lty = 2, size = .25) +
  geom_pointrangeh(
    aes(xmin = q2.5, xmax = q97.5),
    position = position_dodge2v(.25)
  ) +
  facet_wrap("variable") +
  labs(x = "Correlation") +
  theme(axis.title.y = element_blank())
cor_pars_brm %>% 
  mutate(across(where(is.numeric), ~number(., .01))) %>% 
  transmute(
    type, outcome, variable,
    r = str_glue("{mean} [{q2.5}, {q97.5}]")
  ) %>% 
  pivot_wider(names_from = variable, values_from = r) %>% 
  arrange(outcome, type) %>% 
  kbl() %>% 
  kable_custom()
```

### Internet

#### Draw scatterplots

```{r}
get_country_coefs <- function(outcome, predictor, model) {
  x <- readRDS(str_glue("models/brm-{outcome}-{model}.rds"))
  x_country <- coef(x)$country[
    ,,c(
      str_glue("{predictor}_Intercept"), 
      "val_Intercept", 
      str_glue("{predictor}_year"), 
      "val_year"
      )
    ] %>%
    as.data.frame() %>% 
    rownames_to_column("country") %>% 
    tibble() %>% 
    select(country, contains("Estimate"))
  # Below adds region ranefs to country coefs to produce predictions
  x_region <- ranef(x)$region[
    ,,c(
      str_glue("{predictor}_Intercept"), 
      "val_Intercept", 
      str_glue("{predictor}_year"), 
      "val_year"
      )
    ] %>%
    as.data.frame() %>%
    rownames_to_column("region") %>%
    tibble() %>%
    select(region, contains("Estimate"))
  x_country %>%
    left_join(distinct(dat, region, country)) %>%
    left_join(x_region, by = c("region")) %>%
    # Rename X for function
    rename_with(~str_replace(., "internet|mobile", "x")) %>% 
    transmute(
      country, region,
      Estimate.x_Intercept = Estimate.x_Intercept.x + Estimate.x_Intercept.y,
      Estimate.x_year = Estimate.x_year.x + Estimate.x_year.y,
      Estimate.val_Intercept = Estimate.val_Intercept.x + Estimate.val_Intercept.y,
      Estimate.val_year = Estimate.val_year.x + Estimate.val_year.y,
    )
}
get_country_coefs <- memoise(get_country_coefs, cache = cd)
brm_country_coefs_internet <- fits %>%
  mutate(out = map(outcome, ~get_country_coefs(.x, "internet", "1"))) %>% 
  unnest(out)

p_means_brm_internet <- brm_country_coefs_internet %>%
  ggplot(aes(Estimate.x_Intercept, Estimate.val_Intercept, col = region)) +
  scale_color_discreterainbow(
    name = "Region",
    labels = ~str_replace(., "_", " ")
  ) +
  geom_point(
    shape = 1, size = 1.0, alpha = .75
  ) +
  scale_y_continuous(
    labels = ~percent(./100, .1)
  ) +
  facet_wrap(
    ~outcome,
    scales = "free", nrow = 1,
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  theme(
    aspect.ratio = 1
  ) +
  guides(
    color = guide_legend(
      override.aes = list(alpha = 1, shape = 16, size = 1.5)
    )
  ) +
  labs(x = "Internet users per capita (logits)", y = "Outcome")

p_delta_brm_internet <- p_means_brm_internet +
  aes(Estimate.x_year, Estimate.val_year) +
  labs(
    y = "Change in outcome",
    x = "Change in internet users per capita (logits)"
  )
```

### Mobile

#### Draw scatterplots

```{r}
brm_country_coefs_mobile <- fits %>%
  mutate(out = map(outcome, ~get_country_coefs(.x, "mobile", "4"))) %>% 
  unnest(out)

p_means_brm_mobile <- brm_country_coefs_mobile %>%
  ggplot(aes(Estimate.x_Intercept, Estimate.val_Intercept, col = region)) +
  scale_color_discreterainbow(
    name = "Region",
    labels = ~str_replace(., "_", " ")
  ) +
  scale_x_continuous(
    labels = ~percent(., 1),
    breaks = c(-.25, 0, .25, .5, .75, 1)
  ) +
  geom_point(
    shape = 1, size = 1.0, alpha = .75
  ) +
  scale_y_continuous(
    labels = ~percent(./100, .1)
  ) +
  facet_wrap(
    ~outcome,
    scales = "free", nrow = 1, drop = FALSE,
    labeller = as_labeller(~str_replace(., "_", " "))
  ) +
  theme(
    aspect.ratio = 1
  ) +
  guides(
    color = guide_legend(
      override.aes = list(alpha = 1, shape = 16, size = 1.5)
    )
  ) +
  labs(x = "Mobile broadband subscriptions per capita", y = "Outcome")

p_delta_brm_mobile <- p_means_brm_mobile +
  scale_x_continuous(
    labels = ~percent(., 1),
    breaks = extended_breaks(5)
  ) +
  aes(Estimate.x_year, Estimate.val_year) +
  labs(
    y = "Change in outcome",
    x = "Change in mobile broadband subscriptions per capita"
  )
```

```{r}
(
  p_means_ols_internet + ggtitle("A") +
    p_delta_brm_internet + ggtitle("B") +
  p_means_ols_mobile + ggtitle("C") +
    p_delta_brm_mobile + coord_cartesian(xlim = c(0, 2)) + ggtitle("D")
  ) +
  plot_layout(nrow = 4, guides = "collect") & 
  theme(legend.position = "bottom")

agg_png(
  "figures/Fig-country-correlations.png",
  width = W, height = W*0.9, units = "in", res = 400
)
last_plot()
dev.off()
```