### 2022-03-30 02-results.Rmd
Small fixes

# Main results

```{r setup}
#| cache: false
# Packages
library(scales)
library(kableExtra)
library(ggstance)
library(brms)
library(tidybayes)
library(ggdist)
library(memoise)
library(cachem)
library(posterior)
library(ggh4x)
library(parameters)
library(emmeans)
library(patchwork)
library(multidplyr)
library(tidyverse)

cluster <- new_cluster(3)

cd <- cache_disk("memoise_cache", max_size = Inf)

oo <- c(
  "Life_satisfaction", "Negative_experiences", "Positive_experiences",
  "Anxiety", "Depression", "Selfharm"
)

fits <- tibble(outcome = oo) %>%
  mutate(outcome = factor(outcome, levels = oo))

dat <- readRDS("data/data-all.rds")
```

We then turn to the third research question, to what extent are the changes in well-being, and associations between internet adoption and well-being different between the different demographic groups within adolescents? To answer, Models 1 and 2 also included age and sex as moderators of the trends and associations. In this section, we present results based on the moderated effects.

## Time courses

This figure shows the time courses of each of the six outcome variables. Panels by outcome (rows) and region (columns; including World). Each panel shows the time course by all ages and sex, and on average across ages. So we need to calculate four kinds of contrasts: sex\*age | world, sex | world, sex\*age | region, sex | region.

Calculate contrast by sex at the world level.

```{r}
#' Get contrast at average level (no random effects)
#'
#' @param outcome Life_satisfaction etc.
#' @param h Strings representing hypotheses (brm::hypothesis())
#' @param model 1, 2, 3, or 4
get_hyp_avg <- function(outcome, h, model) {
  fit <- readRDS(str_glue("models/brm-{outcome}-{model}.rds"))
  hypothesis(fit, h)$hypothesis
}
get_hyp_avg <- memoise(get_hyp_avg, cache = cd)

h_timecourses <- c(
  Slope_Male = "val_year + val_year:sex1 * -1 = 0",
  Slope_Female = "val_year + val_year:sex1 * 1 = 0"
)

cluster_library(cluster, c("brms", "purrr", "stringr"))
cluster_copy(cluster, c("get_hyp_avg", "h_timecourses"))

timecourses_world_average <- fits %>% 
  group_by(outcome) %>% 
  partition(cluster) %>% 
  mutate(
    out = map(
      outcome,
      ~get_hyp_avg(.x, h = h_timecourses, model = "1")
    )
  ) %>% 
  collect()
```

And then for each sex*age at the world level.

```{r}
#' Get contrast for each age
#'
#' @param outcome Life_satisfaction etc
#' @param h brms hypothesis
#' @param model 1,2,3, or 4
get_hyp_age <- function(outcome, h, model) {
  fit <- readRDS(str_glue("models/brm-{outcome}-{model}.rds"))
  hypothesis(fit, h, scope = "coef", group = "age")$hypothesis
}
get_hyp_age <- memoise(get_hyp_age, cache = cd)
cluster_copy(cluster, "get_hyp_age")

timecourses_world_age <- fits %>% 
  group_by(outcome) %>% 
  partition(cluster) %>% 
  mutate(
    out = map(
      outcome,
      ~get_hyp_age(.x, h = h_timecourses, model = "1")
    )
  ) %>% 
  collect()
```

Combine these to one and plot

```{r}
bind_rows(
  timecourses_world_age %>% 
    unnest(out),
  timecourses_world_average %>% 
    unnest(out) %>% 
    mutate(Group = "Average")
) %>% 
  ungroup() %>% 
  mutate(Group = fct_relevel(Group, "Average", "10 to 14")) %>% 
  select(outcome:CI.Upper) %>% 
  separate(Hypothesis, c("Parameter", "Sex")) %>% 
  mutate(s = sign(CI.Lower) == sign(CI.Upper)) %>% 
  ggplot(aes(Group, Estimate, ymin = CI.Lower, ymax = CI.Upper, fill = Sex, col = Sex, group = Sex)) +
  scale_color_bright() +
  scale_fill_bright() +
  scale_shape_manual(values = c(21, 19), guide = NULL) +
  scale_y_continuous(
    breaks = extended_breaks(),
    # Effect of 10%
    labels = ~percent(./1000), 
    expand = expansion(.4)
  ) +
  geom_hline(yintercept = 0, lty = 2, size = .2) +
  geom_ribbon(
    data = . %>% filter(Group != "Average"),
    aes(), 
    alpha = .2, col = NA
  ) +
  geom_line(
    data = . %>% filter(Group != "Average"),
    aes(group = Sex),
    size = .6
  ) +
  geom_pointrange(
    data = . %>% filter(Group == "Average"),
    position = position_dodge(.75),
    size = .4, fatten = 2.5
  ) +
  facet_wrap(
    "outcome", scales = "free_y", nrow = 2,
    labeller = as_labeller(~str_replace(., "_", " "))) +
  theme(
    axis.text.x = element_text(angle = 90),
    legend.position = "none",
    axis.title.x = element_blank()
  )
```

And for all regions too.

First we calculate the contrast by sex for each region.

```{r}
get_region_timecourse_avg <- function(outcome) {
  x <- readRDS(str_glue("models/brm-{outcome}-1.rds"))
  tmp <- x$data %>% 
    distinct(region, sex) %>% 
    # Note 1 for 10 year change
    crossing(year = c(0, 1), se = 0, itu = TRUE, val = 1) %>% 
    add_epred_draws(
      x,
      re_formula = ~ (year*sex | region),
      resp = "val",
      ndraws = 1000
    ) %>% 
    ungroup()
  tmp %>% 
    select(-.row, -se, -.chain, -.iteration, -val) %>% 
    pivot_wider(names_from = year, values_from = .epred) %>% 
    mutate(delta = `1` - `0`) %>% 
    select(region, sex, .draw, delta) %>% 
    group_by(region, sex) %>% 
    mean_qi(delta)
}
get_region_timecourse_avg <- memoise(
  get_region_timecourse_avg, cache = cd
)

cluster_copy(cluster, "get_region_timecourse_avg")
cluster_library(cluster, c("dplyr", "tidyr", "tidybayes"))
timecourses_region_average <- fits %>% 
  group_by(outcome) %>% 
  partition(cluster) %>% 
  mutate(out = map(outcome, get_region_timecourse_avg)) %>% 
  collect()
```

Then we calculate the contrasts for each sex*age for each region.

```{r}
get_region_timecourse_age <- function(outcome) {
  x <- readRDS(str_glue("models/brm-{outcome}-1.rds"))
  tmp <- x$data %>% 
    distinct(region, sex, age) %>% 
    # Note 1 for decade change
    crossing(year = c(0, 1), se = 0, itu = TRUE, val = 1) %>% 
    add_epred_draws(
      x,
      re_formula = ~ (year*sex | age) + 
        (year*sex | region) + (year*sex | age:region),
      resp = "val",
      ndraws = 1000
    ) %>% 
    ungroup()
  tmp %>% 
    select(-.row, -se, -.chain, -.iteration, -val) %>% 
    pivot_wider(names_from = year, values_from = .epred) %>% 
    mutate(delta = `1` - `0`) %>% 
    select(region, sex, age, .draw, delta) %>% 
    group_by(region, sex, age) %>% 
    mean_qi(delta)
}
get_region_timecourse_age <- memoise(
  get_region_timecourse_age, cache = cd
)

cluster_copy(cluster, "get_region_timecourse_age")
timecourses_region_age <- fits %>% 
  group_by(outcome) %>% 
  partition(cluster) %>% 
  mutate(out = map(outcome, get_region_timecourse_age)) %>% 
  collect()
```

Combine these with world level estimates and plot

```{r}
timecourses_world <- bind_rows(
  timecourses_world_age %>% 
    unnest(out),
  timecourses_world_average %>% 
    unnest(out) %>% 
    mutate(Group = "Average")
) %>% 
  ungroup() %>% 
  separate(Hypothesis, c("Parameter", "sex")) %>% 
  select(
    outcome,
    age = Group,
    sex,
    delta = Estimate,
    .lower = CI.Lower,
    .upper = CI.Upper,
  ) %>% 
  mutate(region = "World")

x <- sort(unique(timecourses_world$age))
x2 <- as.character(x)
x2[str_detect(x2, "5 ")] <- ""

bind_rows(
  timecourses_region_age %>% 
    unnest(out),
  timecourses_region_average %>% 
    unnest(out) %>% 
    mutate(age = "Average")
) %>% 
  ungroup() %>% 
  bind_rows(timecourses_world) %>% 
  mutate(age = fct_relevel(age, "Average", "10 to 14")) %>% 
  mutate(sex = factor(sex, levels = c("Female", "Male"))) %>% 
  arrange(sex) %>% 
  select(outcome:.upper) %>% 
  mutate(s = sign(.lower) == sign(.upper)) %>% 
  ggplot(
    aes(
      age, delta, 
      ymin = .lower, ymax = .upper, 
      fill = sex, col = sex, group = sex
    )
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  scale_shape_manual(values = c(21, 19), guide = NULL) +
  scale_x_discrete(
    breaks = x,
    labels = x2,
    expand = expansion(c(0.02, 0.05))
  ) +
  scale_y_continuous(
    "Estimated linear change per decade",
    breaks = extended_breaks(),
    labels = ~percent(./100),
    expand = expansion(.1)
  ) +
  geom_hline(yintercept = 0, lty = 2, size = .2) +
  geom_ribbon(
    data = . %>% filter(age != "Average"),
    aes(), 
    alpha = .15, col = NA
  ) +
  geom_line(
    data = . %>% filter(age != "Average"),
    aes(group = sex),
    size = .4
  ) +
  geom_pointrange(
    data = . %>% filter(age == "Average"),
    position = position_dodge(.9),
    size = .3, fatten = 0.1
  ) +
  facet_grid(
    outcome~region, scales = "free_y", 
    labeller = as_labeller(~str_replace(., "_", "\n"))) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5),
    legend.position = "none",
    axis.title.x = element_blank()
  )
agg_png(
  "figures/Fig-timecourses-demographics.png", 
  width = W, height = W*0.6, units = "in", res = 400
)
last_plot()
dev.off()
```

## Associations

### Internet

Calculate contrast by sex at the world level.

```{r}
h_internet_associations <- c(
  Slope_Male = "i1_cw + i1_cw:sex1 * -1 = 0",
  Slope_Female = "i1_cw + i1_cw:sex1 * 1 = 0"
)
cluster_copy(cluster, "h_internet_associations")

internet_associations_world_average <- fits %>% 
  group_by(outcome) %>% 
  partition(cluster) %>% 
  mutate(
    out = map(
      outcome,
      ~get_hyp_avg(.x, h = h_internet_associations, model = "2")
    )
  ) %>% 
  collect()
```

And then for each sex*age at the world level.

```{r}
internet_associations_world_age <- fits %>% 
  group_by(outcome) %>% 
  partition(cluster) %>% 
  mutate(
    out = map(
      outcome,
      ~get_hyp_age(.x, h = h_internet_associations, model = "2")
    )
  ) %>% 
  collect()
```

And for all regions too.

First we calculate the contrast by sex for each region.

```{r}
get_internet_region_association_avg <- function(outcome) {
  x <- readRDS(str_glue("models/brm-{outcome}-2.rds"))
  tmp <- x$data %>% 
    distinct(region, sex) %>% 
    # Note .1 for 10% change
    crossing(i1_cw = c(0, .1), se = 0, year = 0) %>% 
    add_epred_draws(
      x,
      re_formula = ~ (i1_cw*sex | region),
      ndraws = 1000
    ) %>% 
    ungroup()
  tmp %>% 
    select(-.row, -se, -.chain, -.iteration) %>% 
    pivot_wider(names_from = i1_cw, values_from = .epred) %>% 
    mutate(delta = `0.1` - `0`) %>% 
    select(region, sex, .draw, delta) %>% 
    group_by(region, sex) %>% 
    mean_qi(delta)
}
get_internet_region_association_avg <- memoise(
  get_internet_region_association_avg, cache = cd
)

cluster_copy(cluster, "get_internet_region_association_avg")
cluster_library(cluster, c("dplyr", "tidyr", "purrr", "stringr", "tidybayes"))
internet_associations_region_average <- fits %>% 
  group_by(outcome) %>% 
  partition(cluster) %>% 
  mutate(out = map(outcome, get_internet_region_association_avg)) %>% 
  collect()
```

#### Demographics

Then we calculate the contrasts for each sex*age for each region.

```{r}
 get_internet_region_association_age <- function(outcome) {
  x <- readRDS(str_glue("models/brm-{outcome}-2.rds"))
  tmp <- x$data %>% 
    distinct(region, sex, age) %>% 
    # Note .1 for 10% change
    crossing(i1_cw = c(0, .1), se = 0, year = 0) %>% 
    add_epred_draws(
      x,
      re_formula = ~ (i1_cw*sex | age) + 
        (i1_cw*sex | region) + (i1_cw*sex | age:region),
      ndraws = 1000
    ) %>% 
    ungroup()
  tmp %>% 
    select(-.row, -se, -.chain, -.iteration) %>% 
    pivot_wider(names_from = i1_cw, values_from = .epred) %>% 
    mutate(delta = `0.1` - `0`) %>% 
    select(region, sex, age, .draw, delta) %>% 
    group_by(region, sex, age) %>% 
    mean_qi(delta)
}
get_internet_region_association_age <- memoise(
  get_internet_region_association_age, cache = cd
)

cluster_copy(cluster, "get_internet_region_association_age")
internet_associations_region_age <- fits %>% 
  group_by(outcome) %>% 
  partition(cluster) %>% 
  mutate(out = map(outcome, get_internet_region_association_age)) %>% 
  collect()
```

Combine these with world level estimates and plot

```{r}
internet_associations_world <- bind_rows(
  internet_associations_world_age %>% 
    unnest(out),
  internet_associations_world_average %>% 
    unnest(out) %>% 
    mutate(Group = "Average")
) %>% 
  ungroup() %>% 
  separate(Hypothesis, c("Parameter", "sex")) %>% 
  mutate(
    # Convert to effect of 10% change in internet/mobile
    delta = Estimate/10,
    .lower = CI.Lower/10,
    .upper = CI.Upper/10
  ) %>% 
  select(
    outcome,
    age = Group,
    sex,
    delta,
    .lower,
    .upper,
  ) %>% 
  mutate(region = "World")

internet_associations <- bind_rows(
  internet_associations_region_age %>% 
    unnest(out),
  internet_associations_region_average %>% 
    unnest(out) %>% 
    mutate(age = "Average")
) %>% 
  ungroup() %>% 
  bind_rows(internet_associations_world) %>% 
  mutate(age = fct_relevel(age, "Average", "10 to 14")) %>% 
  mutate(sex = factor(sex, levels = c("Female", "Male"))) %>% 
  arrange(sex) %>% 
  select(outcome:.upper) %>% 
  mutate(s = sign(.lower) == sign(.upper))
internet_associations %>%   
  ggplot(aes(
    age, delta, 
    ymin = .lower, ymax = .upper, 
    fill = sex, col = sex, group = sex)
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  scale_shape_manual(values = c(21, 19), guide = NULL) +
  scale_x_discrete(
    breaks = x,
    labels = x2,
    expand = expansion(c(0.02, 0.06))
  ) +
  scale_y_continuous(
    "Conditional effect of 10% increase\nin per capita internet users",
    breaks = extended_breaks(),
    labels = ~percent(./100),
    expand = expansion(c(.25, .25))
  ) +
  geom_hline(yintercept = 0, lty = 2, size = .2) +
  geom_ribbon(
    data = . %>% filter(age != "Average"),
    aes(), 
    alpha = .15, col = NA
  ) +
  geom_line(
    data = . %>% filter(age != "Average"),
    aes(group = sex),
    size = .4
  ) +
  geom_pointrange(
    data = . %>% filter(age == "Average"),
    position = position_dodge(.9),
    size = .2, fatten = 1.25
  ) +
  facet_grid(
    outcome~region, scales = "free_y", drop = FALSE,
    labeller = as_labeller(~str_replace(., "_", "\n"))) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5),
    legend.position = "none",
    axis.title.x = element_blank()
  )
agg_png(
  "figures/Fig-associations-internet-demographics.png", 
  width = W, height = W*0.6, units = "in", res = 400
)
last_plot()
dev.off()
```

#### Compare ages

We then compared, for selfharm, the youngest age group's association to all other age group's associations.

```{r}
fit_selfharm <- readRDS("models/brm-Selfharm-3.rds")
h <- hypothesis(fit_selfharm, "m1_cw = 0", scope = "coef", group = "age")
s <- tibble(h$samples)
hypothesis(s, paste0("H1 > H", 2:9))$hypothesis %>% 
  as.data.frame() %>% 
  kbl(digits = 4) %>% 
  kable_custom()
```

#### Forest plot

```{r}
library(ggstance)
internet_associations %>% 
  filter(age == "Average") %>% 
  ggplot(aes(delta, region, col = sex)) +
  geom_vline(xintercept = 0, lty = 2, size = .2) +
  geom_pointrangeh(
    aes(xmin = .lower, xmax = .upper),
    position = position_dodge2v(.3)
  ) +
  facet_wrap("outcome", scales = "free_x", ncol = 3)
```

### Mobile

This figure shows mobile --> outcome associations. Panels by outcome (rows) and region (columns; including World). Each panel shows the association by all ages and sex, and on average across ages. So we need to calculate four kinds of contrasts: sex\*age | world, sex | world, sex\*age | region, sex | region.

Calculate contrast by sex at the world level.

```{r}
get_hyp_avg <- function(outcome, h, model) {
  fit <- readRDS(str_glue("models/brm-{outcome}-{model}.rds"))
  hypothesis(fit, h)$hypothesis
}
get_hyp_avg <- memoise(get_hyp_avg, cache = cd)

h_mobile_associations <- c(
  # Intercept_Male = "Intercept + sex1 * -1 = 0",
  # Intercept_Female = "Intercept + sex1 * 1 = 0",
  Slope_Male = "m1_cw + m1_cw:sex1 * -1 = 0",
  Slope_Female = "m1_cw + m1_cw:sex1 * 1 = 0"
)

cluster_library(cluster, c("brms", "purrr", "stringr"))
cluster_copy(cluster, c("get_hyp_avg", "h_mobile_associations"))

mobile_associations_world_average <- fits %>% 
  group_by(outcome) %>% 
  partition(cluster) %>% 
  mutate(
    out = map(
      outcome,
      ~get_hyp_avg(.x, h = h_mobile_associations, model = "3")
    )
  ) %>% 
  collect()
```

And then for each sex*age at the world level.

```{r}
get_hyp_age <- function(outcome, h, model) {
  fit <- readRDS(str_glue("models/brm-{outcome}-{model}.rds"))
  hypothesis(fit, h, scope = "coef", group = "age")$hypothesis
}
get_hyp_age <- memoise(get_hyp_age, cache = cd)
cluster_copy(cluster, "get_hyp_age")

mobile_associations_world_age <- fits %>% 
  group_by(outcome) %>% 
  partition(cluster) %>% 
  mutate(
    out = map(
      outcome,
      ~get_hyp_age(.x, h = h_mobile_associations, model = "3")
    )
  ) %>% 
  collect()
```

Combine these to one and plot

```{r}
bind_rows(
  mobile_associations_world_age %>% 
    unnest(out),
  mobile_associations_world_average %>% 
    unnest(out) %>% 
    mutate(Group = "Average")
) %>% 
  ungroup() %>% 
  mutate(Group = fct_relevel(Group, "Average", "10 to 14")) %>% 
  select(outcome:CI.Upper) %>% 
  separate(Hypothesis, c("Parameter", "Sex")) %>% 
  mutate(s = sign(CI.Lower) == sign(CI.Upper)) %>% 
  ggplot(aes(Group, Estimate, ymin = CI.Lower, ymax = CI.Upper, fill = Sex, col = Sex, group = Sex)) +
  scale_color_bright() +
  scale_fill_bright() +
  scale_shape_manual(values = c(21, 19), guide = NULL) +
  scale_y_continuous(
    breaks = extended_breaks(),
    # Effect of 10%
    labels = ~percent(./1000), 
    expand = expansion(.4)
  ) +
  geom_hline(yintercept = 0, lty = 2, size = .2) +
  geom_ribbon(
    data = . %>% filter(Group != "Average"),
    aes(), 
    alpha = .2, col = NA
  ) +
  geom_line(
    data = . %>% filter(Group != "Average"),
    aes(group = Sex),
    size = .6
  ) +
  geom_pointrange(
    data = . %>% filter(Group == "Average"),
    position = position_dodge(.75),
    size = .4, fatten = 2.5
  ) +
  facet_wrap(
    "outcome", scales = "free_y", nrow = 2,
    labeller = as_labeller(~str_replace(., "_", " "))) +
  theme(
    axis.text.x = element_text(angle = 90),
    legend.position = "none",
    axis.title.x = element_blank()
  )
```

And for all regions too.

First we calculate the contrast by sex for each region.

```{r}
get_region_association_avg <- function(outcome) {
  x <- readRDS(str_glue("models/brm-{outcome}-3.rds"))
  tmp <- x$data %>% 
    distinct(region, sex) %>% 
    # Note .1 for 10% change
    crossing(m1_cw = c(0, .1), se = 0, year = 0, val = 1) %>% 
    add_epred_draws(
      x,
      re_formula = ~ (m1_cw*sex | region),
      ndraws = 1000
    ) %>% 
    ungroup()
  tmp %>% 
    select(-.row, -se, -.chain, -.iteration, -val) %>% 
    pivot_wider(names_from = m1_cw, values_from = .epred) %>% 
    mutate(delta = `0.1` - `0`) %>% 
    select(region, sex, .draw, delta) %>% 
    group_by(region, sex) %>% 
    mean_qi(delta)
}
get_region_association_avg <- memoise(
  get_region_association_avg, cache = cd
)

cluster_copy(cluster, "get_region_association_avg")
cluster_library(cluster, c("dplyr", "tidyr", "purrr", "stringr", "tidybayes"))
mobile_associations_region_average <- fits %>% 
  group_by(outcome) %>% 
  partition(cluster) %>% 
  mutate(out = map(outcome, get_region_association_avg)) %>% 
  collect()
```

Then we calculate the contrasts for each sex*age for each region.

```{r}
get_region_association_age <- function(outcome) {
  x <- readRDS(str_glue("models/brm-{outcome}-3.rds"))
  tmp <- x$data %>% 
    distinct(region, sex, age) %>% 
    # Note .1 for 10% change
    crossing(m1_cw = c(0, .1), se = 0, year = 0, val = 1) %>% 
    add_epred_draws(
      x,
      re_formula = ~ (m1_cw*sex | age) + 
        (m1_cw*sex | region) + (m1_cw*sex | age:region),
      ndraws = 1000
    ) %>% 
    ungroup()
  tmp %>% 
    select(-.row, -se, -.chain, -.iteration, -val) %>% 
    pivot_wider(names_from = m1_cw, values_from = .epred) %>% 
    mutate(delta = `0.1` - `0`) %>% 
    select(region, sex, age, .draw, delta) %>% 
    group_by(region, sex, age) %>% 
    mean_qi(delta)
}
get_region_association_age <- memoise(
  get_region_association_age, cache = cd
)

cluster_library(cluster, c("dplyr", "tidyr", "tidybayes", "purrr"))
cluster_copy(cluster, "get_region_association_age")
mobile_associations_region_age <- fits %>% 
  group_by(outcome) %>% 
  partition(cluster) %>% 
  mutate(out = map(outcome, get_region_association_age)) %>% 
  collect()
```

Combine these with world level estimates and plot

```{r}
mobile_associations_world <- bind_rows(
  mobile_associations_world_age %>% 
    unnest(out),
  mobile_associations_world_average %>% 
    unnest(out) %>% 
    mutate(Group = "Average")
) %>% 
  ungroup() %>% 
  separate(Hypothesis, c("Parameter", "sex")) %>% 
  mutate(
    # Convert to effect of 10% change in internet/mobile
    delta = Estimate/10,
    .lower = CI.Lower/10,
    .upper = CI.Upper/10
  ) %>% 
  select(
    outcome,
    age = Group,
    sex,
    delta,
    .lower,
    .upper,
  ) %>% 
  mutate(region = "World")

mobile_associations <- bind_rows(
  mobile_associations_region_age %>% 
    unnest(out),
  mobile_associations_region_average %>% 
    unnest(out) %>% 
    mutate(age = "Average")
) %>% 
  ungroup() %>% 
  bind_rows(mobile_associations_world) %>% 
  mutate(age = fct_relevel(age, "Average", "10 to 14")) %>% 
  mutate(sex = factor(sex, levels = c("Female", "Male"))) %>% 
  arrange(sex) %>% 
  select(outcome:.upper) %>% 
  mutate(s = sign(.lower) == sign(.upper))
mobile_associations %>%   
  ggplot(
    aes(
      age, delta, 
      ymin = .lower, ymax = .upper, 
      fill = sex, col = sex, group = sex
    )
  ) +
  scale_color_bright() +
  scale_fill_bright() +
  scale_shape_manual(values = c(21, 19), guide = NULL) +
  scale_x_discrete(
    breaks = x,
    labels = x2,
    expand = expansion(c(0.02, 0.06))
  ) +
  scale_y_continuous(
    "Conditional effect of 10% increase\nin per capita mobile broadband subscriptions",
    breaks = extended_breaks(),
    labels = ~percent(./100),
    expand = expansion(.4)
  ) +
  geom_hline(yintercept = 0, lty = 2, size = .2) +
  geom_ribbon(
    data = . %>% filter(age != "Average"),
    aes(), 
    alpha = .15, col = NA
  ) +
  geom_line(
    data = . %>% filter(age != "Average"),
    aes(group = sex),
    size = .4
  ) +
  geom_pointrange(
    data = . %>% filter(age == "Average"),
    position = position_dodge(.9),
    size = .2, fatten = 1.25
  ) +
  facet_grid(
    outcome~region, scales = "free_y", 
    labeller = as_labeller(~str_replace(., "_", "\n"))) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5),
    legend.position = "none",
    axis.title.x = element_blank()
  )
agg_png(
  "figures/Fig-associations-mobile-demographics.png", 
  width = W, height = W*0.6, units = "in", res = 400
)
last_plot()
dev.off()
```

#### Forest plot

```{r}
mobile_associations %>% 
  filter(age == "Average") %>% 
  ggplot(aes(delta, region, col = sex)) +
  geom_vline(xintercept = 0, lty = 2, size = .2) +
  geom_pointrangeh(
    aes(xmin = .lower, xmax = .upper),
    position = position_dodge2v(.3)
  ) +
  facet_wrap("outcome", scales = "free_x", ncol = 3)
```