### 2022-04-08 00-describe.Rmd

# Describe

```{r setup}
#| cache: false
# Packages
library(scales)
library(ggtext)
library(lubridate)
library(kableExtra)
library(ggstance)
library(ggh4x)
library(mgcv)
library(tidyverse)

# Load data sets
itu <- read_rds("data/itu.rds")
gwp <- read_rds("data/gwp.rds")
gbd <- read_rds("data/gbd.rds")
d <- read_rds("data/data-all.rds")

oo <- c(
  "Life_satisfaction", "Negative_experiences", "Positive_experiences",
  "Anxiety", "Depression", "Selfharm"
)
```

Number of countries/years/people per outcome

```{r}
bind_rows(
  "GWP" = gwp,
  "GBD" = gbd,
  .id = "Dataset"
) %>% 
  mutate(Dataset = fct_inorder(Dataset)) %>% 
  group_by(Dataset, region, outcome, country) %>% 
  summarise(
    n_n = sum(n),
    n_year = length(unique(year))
  ) %>% 
  summarise(
    N = sum(n_n), 
    Years = mean(n_year)
  ) %>%
  summarise(
    N_sum = round(mean(N)), 
    Years_mean = round(mean(Years))
  ) %>%
  kbl(caption = "Dataset summary") %>% 
  kable_custom()
tmp <- gwp %>% 
  group_by(outcome) %>% 
  summarise(n=sum(n)) %>% 
  summarise(N=comma(median(n))) %>% 
  pull(N)
```

The median number of respondents per (well-being) outcome was `r tmp`, and the unique age bins in the GWP data were `r unique(gwp$age)`.

Number of data points per outcome

```{r}
bind_rows(
  "GWP" = gwp,
  "GBD" = gbd,
  .id = "Dataset"
) %>% 
  mutate(Dataset = fct_inorder(Dataset)) %>% 
  count(Dataset, outcome) %>% 
  kbl(caption = "Observations per outcome") %>% 
  kable_custom()

```

Number of ITU observations

```{r}
tmp <- itu %>% 
  select(region, country, year, internet, mobile)
tmp %>% 
  drop_na(mobile)
```

## Data figure

```{r trends-regions}
#| echo: false
#| fig-width: 12
#| fig-height: 8
#| fig-cap: "Trends in all variables across regions"
#| label: fig-trends-regions

# Stack data sets
# Assume all outcomes are proportions (0-1)
# Draw small multiples figure, outcomes x regions
d <- bind_rows(
  read_rds("data/gwp.rds"),
  read_rds("data/gbd.rds")
)

itu <- read_rds("data/itu.rds") %>%
  select(
    region, country, year,
    Internet = internet,
    Mobile = mobile
  ) %>%
  # Truncate mobile percentages for figure
  mutate(Mobile = if_else(Mobile > 1.5, NaN, Mobile)) %>%
  pivot_longer(
    c(Internet, Mobile),
    names_to = "outcome", values_to = "val"
  )

tmp_country <- bind_rows(itu, d) %>%
  mutate(outcome = fct_inorder(str_replace(outcome, "_", "_"))) %>%
  group_by(region, country, year, outcome) %>%
  summarise(mean = mean(val, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(outcome = factor(outcome, levels = c("Internet", "Mobile", oo)))

tmp_region <- tmp_country %>%
  group_by(region, outcome) %>%
  group_modify(~broom::augment(gam(mean ~ s(year, bs = "cs", k = 5), data = .))) %>%
  ungroup() %>% 
  mutate(outcome = factor(outcome, levels = c("Internet", "Mobile", oo)))

tmp_world <- tmp_country %>%
  group_by(outcome) %>%
  group_modify(~broom::augment(gam(mean ~ s(year, bs = "cs", k = 5), data = .))) %>%
  ungroup() %>%
  mutate(region = "World") %>% 
  mutate(outcome = factor(outcome, levels = c("Internet", "Mobile", oo)))

tmp_country %>%
  mutate(year = make_date(year)) %>%
  ggplot(aes(year, mean)) +
  scale_y_continuous(
    "Value",
    labels = ~percent(., .1),
    breaks = pretty_breaks(),
    expand = expansion(.05)
  ) +
  scale_x_date(
    "Year",
    breaks = make_date(c(2005, 2010, 2015, 2020)),
    date_labels = "'%y",
    expand = expansion(.02)
  ) +
  geom_line(
    aes(group = interaction(country, outcome)),
    alpha = .25, size = .4, color = "dodgerblue1"
  ) +
  geom_line(
    data = tmp_region %>% mutate(year = make_date(year)),
    aes(y = .fitted, group = interaction(region, outcome)),
    alpha = 1, size = .6, color = "dodgerblue4"
  ) +
  geom_line(
    data = tmp_world %>% mutate(year = make_date(year)),
    aes(y = .fitted, group = interaction(region, outcome)),
    alpha = 1, size = .8, color = "dodgerblue4"
  ) +
  facet_grid(
    outcome ~ region,
    scales = "free_y",
    labeller = as_labeller(function(x) str_replace(x, "_", "\n"))
  ) +
  theme(
    aspect.ratio = 1,
    legend.position = "none",
    panel.spacing.x = unit(4, "pt"),
    panel.spacing.y = unit(6, "pt"),
    axis.title.y = element_blank()
  )
agg_png(
  "figures/Fig-trends-data.png",
  width = W, height = W*0.8, units = "in", res = 400
)
last_plot()
dev.off()
```