### 2022-04-08 02-controls.Rmd
Init control analyses

# Control analyses

```{r setup}
#| cache: false
# Packages
library(scales)
library(kableExtra)
library(ggstance)
library(brms)
library(tidybayes)
library(ggdist)
library(memoise)
library(cachem)
library(lubridate)
library(posterior)
library(ggh4x)
library(parameters)
library(janitor)
library(countrycode)
library(emmeans)
library(patchwork)
library(multidplyr)
library(tidyverse)

cluster <- new_cluster(3)

cd <- cache_disk("memoise_cache", max_size = Inf)

oo <- c(
  "Life_satisfaction", "Negative_experiences", "Positive_experiences",
  "Anxiety", "Depression", "Selfharm"
)

fits <- tibble(outcome = oo) %>%
  mutate(outcome = factor(outcome, levels = oo))

dat <- readRDS("data/data-all.rds")
```

Here, we analyse associations between other country-level indicators and well-being and mental health.