### 2022-05-16 index.Rmd
Cleaning

# Preface {.unnumbered}

Welcome! This document is the online analysis supplement to *Global mental health and well-being in the internet age* (Vuorre & Przybylski, [submitted]()). It contains all the analyses reported in the manuscript along with supplemental figures and data. The data and source code are on [GitHub](https://github.com/digital-wellbeing/global-wbmh), and an archived permanent copy of these are on [OSF](https://osf.io/ys7m9/).

### How to reproduce our results

The analyses were conducted in R, and can be reproduced locally:

1. Clone the github repo
  - Terminal: `git clone https://github.com/digital-wellbeing/global-wbmh.git`  
      OR
  - RStudio: File -> New Project -> Version Control -> Git -> use the URL from above

2. Prepare the R environment
  - Terminal: `Rscript -e "renv::restore()"`  
      OR
  - RStudio: Click renv -> Restore Library in the Packages panel

3. Render the source files to a Quarto book
 - Install [Quarto](https://quarto.org/docs/getting-started/installation.html) if you don't already have it
  - Terminal: `quarto render`  
      OR
  - RStudio: Open the project (`global-wbmh.Rproj`) and click Render in the Build panel

If you encounter problems, please [open an issue](https://github.com/digital-wellbeing/global-wbmh/issues).

::: {.callout-important}
The project repo includes the GBD and ITU datasets, and a synthetic mock version of the GWP dataset to enable reproducing all our computations.
**The models take several hours/days each to run**---depending on your local computing resources---and therefore the rendering process can take several days. For this reason, the build will fail after having cleaned the data. Then, run `models.R` with settings specific to your environment/cluster. Once that is done you can build again and it should work.
:::