### 2022-10-26 ms.Rmd
Major revision

---
title: Global well-being and mental health in the internet age
shorttitle: Global well-being
leftheader: Global well-being
author: 
  - name: Matti Vuorre
    affiliation: 1
    address: Oxford Internet Institute, University of Oxford
    email: matti.vuorre@oii.ox.ac.uk
  - name: Andrew K. Przybylski
    affiliation: 1
    corresponding: yes
    address: Oxford Internet Institute, University of Oxford
    email: andy.przybylski@oii.ox.ac.uk
affiliation:
  - id: 1
    institution: Oxford Internet Institute, University of Oxford
authornote: |
  \noindent \textbf{This pre-print is not yet peer-reviewed.}
abstract: | 
  In the last two decades the rise of internet technologies have impacted nearly all domains of human life and inspired concern that their widespread adoption has impacted mental health and psychological well-being. However, research on the topic is contested and hampered by methodological shortcomings leaving the broader consequences of internet adoption unknown. We show that the past two decades have seen small and inconsistent changes in global well-being and mental health. The broadest available health and internet technology data does not support the idea that the adoption of internet and mobile broadband is consistently linked to negative psychological outcomes. However, associations between internet technology adoption and depression and self-harm were strongest for individuals aged between 10 and 34. Further investigation of this topic requires transparent study of online behaviours where they occur, namely on online platforms. We call for increased collaborative efforts between independent scientists and the internet technology sector.

keywords: well-being, mental health, internet technology, technology effects
wordcount: '`r scales::number(wordcountaddin:::word_count("ms.Rmd"), big.mark = ",")`'
bibliography: "references.bib"
floatsintext: yes
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
# header-includes:
csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass: apa7
classoption: jou
output: 
  papaja::apa6_pdf:
    number_sections: false
    keep_tex: false
  papaja::apa6_docx: 
    number_sections: false
editor_options: 
  chunk_output_type: console
always_allow_html: true
---

```{r setup, include = FALSE}
# Packages
library(papaja)
library(readxl)
library(imputeTS)
library(scales)
library(janitor)
library(haven)
library(labelled)
library(tidybayes)
library(ggdist)
library(countrycode)
library(lubridate)
library(brms)
library(patchwork)
library(knitr)
library(ggstance)
library(bayestestR)
library(ggh4x)
library(tidyverse)

opts_chunk$set(
  eval = TRUE,
  cache = TRUE,
  include = FALSE,
  message = FALSE,
  warning = FALSE,
  dpi = 300,
  fig.align = "center"
)
# For tables: Show NAs as blanks
options(knitr.kable.NA = '')

# Create directory for intermediate files
dir.create("data", FALSE)

# Plotting options
theme_set(
  theme_linedraw(
    base_size = if_else(knitr::is_latex_output(), 9, 8)
  ) +
    theme(
      panel.grid = element_blank(),
      strip.text = element_text(margin = margin(4, 4, 4, 4, "pt")),
      axis.text = element_text(size = rel(0.75)),
      axis.ticks = element_line(size = rel(0.3))
    )
)
```

```{r functions}
# Function to replace extreme percentages
percent2 <- function(x, accuracy = .1) {
  x <- percent(x, accuracy = accuracy)
  x <- if_else(x == "100.0%", ">99.9%", x)
  x <- if_else(x == "0.0%", "<0.1%", x)
  x
}
```

```{r data-clean}
## # Understand variables & values
## # Codes NOTE 1: yes 2: no!
## read_spss(
##   "data-raw/Gallup/The_Gallup_101521.sav",
##   n_max = 1,
##   col_select = c(
##     WPID, WP1219,
##     WP16,
##     WP60, WP61, WP63, WP65, WP67,
##     WP68, WP69, WP70, WP71, WP74
##   )
## ) %>%
##   generate_dictionary()

# First liberate the data from the slow SPSS file and save to disk
gwp_path <- "data-raw/Gallup/gwp-processed.rds"
if (!file.exists(gwp_path)) {
  gwp <- read_spss(
    "data-raw/Gallup/The_Gallup_101521.sav",
    col_select = c(
      YEAR_CALENDAR,
      COUNTRYNEW,
      WPID, WP1220, WP1219,
      WP16,
      WP60, WP61, WP63, WP65, WP67,
      WP68, WP69, WP70, WP71, WP74
    )
  )
  # Get rid of SPSS attributes
  gwp <- gwp %>%
    zap_labels() %>%
    zap_label() %>%
    zap_widths() %>%
    zap_formats()
  # Save into a good format
  write_rds(gwp, gwp_path)
} else {
  gwp <- read_rds(gwp_path)
}

# Rename and recode variables
gwp <- gwp %>%
  clean_names() %>%
  transmute(
    country = countrynew,
    year = year_calendar,
    id = wpid,
    sex = factor(wp1219, levels = c(2, 1), labels = c("Female", "Male")),
    age = wp1220,
    # Don't know, refused, and missing values
    Life_satisfaction = if_else(between(wp16, 0, 10), wp16, NaN),
    across(wp60:wp74, ~ if_else(between(., 1, 2), ., NaN)),
    # Also reverse the weird 1: yes 2: no coding here
    across(wp60:wp74, ~ 3 - .)
  )

# Scale scores note rescaling to percentages
gwp <- gwp %>%
  mutate(
    Life_satisfaction = Life_satisfaction * 10,
    Negative_experiences =
      (rowMeans(select(., wp68:wp74), na.rm = TRUE) - 1) * 100,
    Positive_experiences =
      (rowMeans(select(., wp60:wp67), na.rm = TRUE) - 1) * 100
  ) %>%
  select(-c(wp60:wp74))

# Categorize ages
gwp <- filter(gwp, age <= 89)
gwp <- gwp %>%
  mutate(
    age = cut(
      age,
      breaks = seq(15, 90, by = 5),
      include.lowest = TRUE,
      right = FALSE,
      labels = paste(seq(15, 85, by = 5), "to", seq(19, 89, by = 5))
    ) %>% as.character()
  )

# There is one 13 year old which resulted in a NA and we drop it
gwp <- drop_na(gwp, age)

# Some sex values are missing, drop those
gwp <- drop_na(gwp, sex)

gwp <- gwp %>%
  pivot_longer(
    c(
      Life_satisfaction,
      Negative_experiences,
      Positive_experiences
    ),
    names_to = "outcome", values_to = "val"
  ) %>%
  group_by(country, year, sex, age, outcome) %>%
  summarise(
    n = n(),
    se = sd(val, na.rm = TRUE) / sqrt(n),
    val = mean(val, na.rm = TRUE)
  ) %>%
  ungroup()

# Note that if there is no response variance, SE will be zero, and if there is only 1 response, SE will be NA. Here we fix those by assigning them the maximum SE (0.5)
gwp <- gwp %>% 
  mutate(
    se = if_else(is.na(se) | se == 0, 0.5, se)
  )

# We can see that this would result in North Cyprus being lumped with Cyprus so we need to replace the name before harmonising

# Harmonise old names and replace only if harmonised name found
gwp <- gwp %>%
  mutate(
    country_harmonised = countryname(
      country,
      destination = "cldr.short.en"
    )
  ) %>%
  # This picks the harmonised name if exists, otherwise original name
  # Prevent north cyprus from becoming cyprus
  mutate(
    country_harmonised = ifelse(
      country == "Northern Cyprus",
      "Northern Cyprus",
      country_harmonised
    )
  ) %>%
  mutate(country = coalesce(country_harmonised, country)) %>%
  select(-country_harmonised)

gwp_mock_path <- "data-raw/gwp-MOCK.rds"
if (dir.exists("data-raw/Gallup") & !file.exists(gwp_mock_path)) {
  library(synthpop)
  gwp_syn <- gwp %>% 
    drop_na()
  gwp_syn <- syn(
    gwp_syn, 
    method = c("", "", "", "", "", "", "normrank", "normrank"),
    maxfaclevels = 168
  )
  gwp_syn <- tibble(gwp_syn$syn)
  saveRDS(gwp_syn, gwp_mock_path)
}

# Load if actual data doesn't exist
if (!dir.exists("data-raw/Gallup")) {
  cat("Using synthetic GWP data.")
  gwp <- read_rds(gwp_mock_path)
} else {
  cat("Using actual GWP data.")
}

# Load data files and merge to one table
gbd <- list.files(
  "data-raw/GBD/",
  pattern = ".zip", full.names = TRUE, recursive = TRUE
) %>%
  read_csv()

# First step is to clean names and remove unnecessary variables
gbd <- clean_names(gbd)
# distinct(gbd, measure, metric)
gbd <- select(gbd, -measure, -metric)

# Rates to percentages
gbd <- gbd %>%
  mutate(across(c(val, upper, lower), ~ .x / 1000))

# Clean cause names
# distinct(gbd, cause)
gbd <- gbd %>%
  mutate(
    cause = case_when(
      cause == "Anxiety disorders" ~ "Anxiety",
      cause == "Depressive disorders" ~ "Depression",
      cause == "Self-harm" ~ "Selfharm"
    )
  )

# Harmonise variable names with other datasets
gbd <- gbd %>%
  rename(country = location, outcome = cause)

# The outcomes values are model predictions and come with lower and upper CI limits (2.5 and 97.5 %iles of their posterior distributions). We convert those to normal approximate standard errors.
gbd <- gbd %>%
  mutate(se = (upper - lower) / (1.96 * 2)) %>%
  select(-c(upper, lower))

# Harmonise old names and replace only if harmonised name found
gbd <- gbd %>%
  mutate(
    country_harmonised = countryname(
      country,
      destination = "cldr.short.en"
    )
  ) %>%
  # This picks the harmonised name if exists, otherwise original name
  mutate(country = coalesce(country_harmonised, country)) %>%
  select(-country_harmonised)

# Download data sets from the ITU website if not yet downloaded
if (!file.exists("data-raw/ITU/PercentIndividualsUsingInternet.xlsx")) {
  dir.create("data-raw/ITU")
  download.file(
    "https://www.itu.int/en/ITU-D/Statistics/Documents/statistics/2022/July/PercentIndividualsUsingInternet.xlsx",
    "data-raw/ITU/PercentIndividualsUsingInternet.xlsx"
  )
  download.file(
    "https://www.itu.int/en/ITU-D/Statistics/Documents/statistics/2022/July/MobileBroadbandSubscriptions_2007-2021.xlsx",
    "data-raw/ITU/MobileBroadbandSubscriptions_2007-2020.xlsx"
  )
}

# Read and reshape mobile broadband data
itu_mobile <- read_xlsx(
  "data-raw/ITU/MobileBroadbandSubscriptions_2007-2020.xlsx",
  sheet = 2,
  col_types = "text"
) %>%
  pivot_longer(-c(Indicator, Economy)) %>%
  separate(name, into = c("year", "variable")) %>%
  pivot_wider(names_from = variable) %>%
  mutate(mobile = as.numeric(value)) %>%
  select(country = Economy, year, mobile)

# Percent using internet
itu_percent <- read_xlsx(
  "data-raw/ITU/PercentIndividualsUsingInternet.xlsx",
  col_types = "text"
) %>%
  pivot_longer(-c(Indicator, Economy)) %>%
  separate(name, into = c("year", "variable")) %>%
  pivot_wider(names_from = variable) %>%
  mutate(internet = as.numeric(value)) %>%
  select(country = Economy, year, internet)

# Merge to one table in wide format (oldest on left to include all years)
itu <- reduce(list(itu_percent, itu_mobile), left_join)

# years to numbers
itu <- itu %>%
  mutate(year = as.numeric(year))

# Drop tables
rm(itu_mobile, itu_percent)

# Harmonise old names and replace only if harmonised name found
itu <- itu %>%
  mutate(
    country_harmonised = countryname(
      country,
      destination = "cldr.short.en"
    )
  ) %>%
  # This picks the harmonised name if exists, otherwise original name
  mutate(country = coalesce(country_harmonised, country)) %>%
  select(-country_harmonised)

# Drop countries that don't exist in outcomes
itu <- itu %>%
  filter(
    country %in% unique(gbd$country) | country %in% unique(gwp$country)
  )

# Drop countries that don't have any actual internet data
itu_zeros <- itu %>%
  group_by(country) %>%
  summarise(s = sum(internet, na.rm = TRUE)) %>%
  filter(s == 0)

itu <- itu %>%
  filter(!(country %in% itu_zeros$country))
rm(itu_zeros)

# Reshape to easily impute all variables
itu_long <- itu %>%
  pivot_longer(c(internet, mobile))

# Impute by country and variable
itu_long <- itu_long %>%
  group_by(country, name) %>%
  # Find first and last year with this variable in each country
  # so as to prevent extrapolation
  mutate(
    min_year = year[min(which(!is.na(value)))],
    max_year = year[max(which(!is.na(value)))]
  ) %>%
  # Interpolate values linearly
  # use possibly() to avoid errors for countries with no data
  mutate(
    across(
      value,
      .fns = list(
        i = possibly(
          function(x) na_interpolation(x, maxgap = Inf),
          otherwise = NaN
        )
      )
    )
  ) %>%
  # Take out projected (non-intermediate) imputed values
  mutate(
    value_i = if_else(
      between(year, unique(min_year), unique(max_year)),
      value_i,
      NaN
    )
  ) %>%
  ungroup() %>%
  select(-min_year, -max_year)

# Summary
itu_sum <- itu_long %>%
  filter(!(name=="mobile" & year < 2007)) %>% 
  group_by(Metric = name) %>%
  summarise(
    total = n(), 
    total_not_missing = sum(!is.na(value)),
    missing = sum(is.na(value)),
    imputed = sum(is.na(value) & !is.na(value_i))
  )

# We then replace the original values with the imputed ones
itu <- itu_long %>%
  # Pick non-missing value, from original and imputed
  mutate(value = coalesce(value, value_i)) %>%
  select(-value_i) %>%
  pivot_wider()
rm(itu_long)

regions <- bind_rows(
  distinct(itu, country),
  distinct(gbd, country),
  distinct(gwp, country)
) %>% 
  distinct() %>% 
  mutate(
    region = countrycode(
      country, 
      origin = "country.name", 
      destination = "continent"
    )
  )

# Manually fix two countries
regions <- regions %>% 
  mutate(
    region = if_else(country == "Kosovo", "Europe", region),
    region = if_else(country == "Nagorno Karabakh", "Asia", region)
  )

# Add regions to data tables
itu <- left_join(itu, regions)
gbd <- left_join(gbd, regions)
gwp <- left_join(gwp, regions)

# Stack outcome data
dat <- bind_rows(gwp, gbd)

# Merge predictors to outcomes
dat <- left_join(dat, itu)

# Center year
dat <- dat %>%
  mutate(year = (year - 2010))

# Ensure that categorical variables are factors
dat <- dat %>%
  mutate(across(c(sex, age), factor))

# Arrange on outcome
oo <- c(
  "Life_satisfaction", "Negative_experiences", "Positive_experiences",
  "Anxiety", "Depression", "Selfharm"
)
dat <- dat %>%
  mutate(outcome = factor(outcome, levels = oo)) %>% 
  arrange(outcome, country, region, year)

# Selfharm models don't converge with SEs
dat <- dat %>% 
  mutate(se = if_else(outcome == "Selfharm", 0, se))

# Save data for models
saveRDS(dat, "data/data-all.rds")
```

In 2005, an estimated 17% of the global population used the internet, but by 2020 this number was already 59% [@ituITUICTStatistics2021]. Accompanying the rapid spread of the internet, worries have proliferated that its broad adoption—and technologies enabled by it such as online games, smartphones, and social media—is actively harming its users, particularly adolescents [@carrShallowsHowInternet2010; @turkleAloneTogetherWhy2011]. As a response, users and a growing number of national governments have acted in limiting access to online technologies [@buckleyChinaTightensLimits2021; @departmentfordigitalculturemedia&sportDraftOnlineSafety2021; @kattulaGamePoliciesContexts2021].

However, evidence for widespread harms of online technologies is limited. Initial reports of internet-facilitated harms have been challenged by later work informed by many methodologies including longitudinal [@jensenYoungAdolescentsDigital2019], specification curve analyses [@orbenAssociationAdolescentWellbeing2019], meta-analyses [@appelAreSocialMedia2020], and systematic reviews [@bestOnlineCommunicationSocial2014; @dickson2018screen; @odgersAnnualResearchReview2020; @ophirNewMediaScreenTime2020]. Reviews indicate early research has been hampered by inaccurate measurements of engagement with internet and related technologies [@davidsonFuzzyConstructsTechnology2022; @parrySystematicReviewMetaanalysis2021; @scharkowAccuracySelfReportedInternet2016], biased convenience samples drawn predominantly from countries in the Global North [@ghaiLackSampleDiversity2021], studying a limited range of well-being outcomes [@orbenSocialMediaEnduring2019], and reliance on self-reported evaluations in place of clinical estimates of important mental health outcomes [@campbellInternetUseSocially2006]. A comprehensive test of the overall association between internet adoption and well-being and mental health, broadly defined, has therefore not been conducted. As a consequence, much of the evidence purporting to show that internet technology adoption is associated with broad changes in well-being and mental health, or even causes negative outcomes, remains equivocal [@dickson2018screen; @hawkesCMOReportUnable2019].

In this research we present two studies of global well-being and mental health in the internet age. In the first, we focus on three aspects of psychological well-being, and contrast them with yearly per-capita internet users and mobile broadband subscriptions, across 168 countries and 16 years. In the second we examine three mental health outcomes across 202 countries and 19 years. Our aim is to better understand: 1) How well-being and mental health have changed, on a global scale, during the past two decades of dramatic proliferation of internet technologies and connectivity; 2) How per-capita internet users and mobile broadband subscriptions predict country-level well-being and mental health within a given country and across countries; and 3) Assess the extent to which associations between internet technology adoption and well-being and mental health differ across age and sex and if they are specific to previously suggested vulnerable populations, such as young women.

# Study 1: Well-being

```{r gwp-data-summaries}
gwp_sum <- gwp %>% 
  group_by(outcome) %>% 
  summarise(n = sum(n), cells = n()) %>% 
  mutate(across(n:cells, ~number(., big.mark = ",")))
```

In Study 1, we focused on psychological well-being, as reflected in self-reports of life satisfaction, positive experiences, and negative experiences from `r gwp_sum$n[1]` individuals aged 15 to 89 across `r length(unique(gwp$country))` countries from `r paste0(range(gwp$year), collapse = " to ")`. We contrasted those observations with time series data of the countries' per capita internet users and mobile broadband subscriptions [@ituITUICTStatistics2021] to examine if and how internet and mobile broadband adoption predicted psychological well-being over the past two decades.

## Transparency and Openness

This study was not preregistered. All code used in this project, along with the GBD data, link to the ITU data, and a synthetic version of the GWP data are available at <https://doi.org/10.5281/zenodo.7004053>. Our sample size was determined by the number of available years, countries, and demographics in each respective dataset.

## Methods

### Internet adoption

We identified the International Telecommunication Union's (ITU) database of information and communications technology (ICT) as the most comprehensive country-level source of time-series data on internet adoption [@ituITUICTStatistics2021]. The ITU has collated, from 222 countries' statistical and telecommunications agencies, the yearly percentages of population using the internet from 2000 to `r max(itu$year)`, and yearly per capita broadband subscriptions from 2007 to `r max(itu$year)`. Our analyses used `r number(itu_sum[1, 3, drop = TRUE], big.mark = ",")` yearly internet percentages and `r number(itu_sum[2, 3, drop = TRUE], big.mark = ",")` broadband subscription rates; we imputed `r number(itu_sum[1, 5, drop = TRUE], big.mark = ",")` (`r number(itu_sum[2, 5, drop = TRUE], big.mark = ",")` mobile) intermediate missing values to this dataset using linear interpolation without extrapolation.

### Subjective well-being

We examined subjective well-being indicators from the Gallup World Poll (GWP), a nationally representative annual survey of 1,000 civilian, non-institutionalised individuals aged 15 years or older from `r length(unique(gwp$country))` countries from 2005 to 2021. The surveys are conducted face-to-face or via telephone, in the respondents' native language and by local interviewers. For details on the GWP sampling and survey methodology, see [@gallupGallupWorldPoll2020; @gallupHowDoesGallup2014].

GWP measures subjective well-being with the positive and negative experience indices, which measure respondents' experienced well-being on the day before the survey with five items each. For positive experiences, these items are "Did you feel well-rested yesterday?"; "Were you treated with respect all day yesterday?"; "Did you smile or laugh a lot yesterday?"; "Did you learn or do something interesting yesterday?"; and "(Did you experience the following feelings during a lot of the day yesterday?) How about enjoyment?" And for negative experiences, the items are responses to "Did you experience the following feelings during a lot of the day yesterday?" with prompts "How about physical pain?"; "How about worry?"; "How about sadness?"; "How about stress?"; and "How about anger?" We aggregated both scales for each respondent by taking a mean of the five items.

Life satisfaction in the moment was measured with one 11-step Likert item, "Please imagine a ladder, with steps numbered from 0 at the bottom to 10 at the top. The top of the ladder represents the best possible life for you and the bottom of the ladder represents the worst possible life for you. On which step of the ladder would you say you personally feel you stand at this time?", similar to the Cantril self-anchoring scale [@cantril1965pattern; @kapteynDimensionsSubjectiveWellbeing2015]. For analyses, we converted these variables to percentages, and aggregated the `r gwp_sum[1, 2]` individuals' data to means and standard errors for each outcome, country, year, sex, and age (5-year age groups from 15 to 89 years) combination (`r gwp_sum[1, 3]` cells).

### Data analysis

We analysed the data with meta-analytic bayesian hierarchical regression models [@burknerBrmsPackageBayesian2017; @gelmanDataAnalysisUsing2007; @standevelopmentteamStanModelingLanguage2021]. We first asked how the well-being outcomes had changed over time. To answer, we modelled each outcome (e.g. life satisfaction) on population level intercepts (we centred time on 2010), contrasts of time (in continuous years), sex (Female: 1, Male: -1), and their interaction. We allowed all coefficients to vary across countries, age groups, and the age by country interactions, and treated all outcomes as normally distributed.

Before modelling the data, we aggregated the individual-level well-being values to means and standard errors for each country-year-age-sex combination, in order to facilitate the Hamiltonian Monte Carlo computations on large data, and to harmonise the analyses and inferences between Studies 1 and 2. Importantly, we note that because we modelled unweighted demographic- and country-level data, our unit of analyses are demographic groups within countries, with no adjustment to e.g. countries' population sizes.

Our second research question asked how changes in internet technology adoption predicted well-being. To answer, we expanded the above model to include within-country centred per capita internet users (or mobile broadband subscriptions in a separate model) and its interaction with sex as predictors of the well-being outcome, and all their associated varying effects. We also included the between-country centered variable as a predictor to examine between-country associations between internet technology adoption and well-being.

Importantly, we did not make the necessary strong assumptions required for identifying causal effects, and therefore highlight the descriptive nature of the resulting associations. For instance, we did not adjust for either time-invariant or -varying confounders, which could bias observed between-country and within-country associations, respectively. Nevertheless, we included time as a predictor to adjust for simple temporal trends. In addition, we within-country centred the internet technology adoption variables in order to isolate within-country associations from between-country associations.

We specified the model as

\begin{align*} 
y_i &\sim \text{Normal}(\mu_i, \sqrt{\sigma^{2} + \text{se}^2_i}), \\
\mu_i &= \alpha_{0} + \beta_{0\text{country}[i]} + \gamma_{0\text{age}[i]} + \delta_{0\text{age:country}[i]} + \\
&\quad\ (\alpha_{1} + \beta_{1\text{country}[i]} + \gamma_{1\text{age}[i]} + \delta_{1\text{age:country}[i]})\text{Time}_i + \\
&\quad\ (\alpha_{2} + \beta_{2\text{country}[i]} + \gamma_{2\text{age}[i]} + \delta_{2\text{age:country}[i]})\text{Sex}_i + \\
&\quad\ (\alpha_{3} + \beta_{3\text{country}[i]} + \gamma_{3\text{age}[i]} + \delta_{3\text{age:country}[i]})\text{Sex}_i \times \text{Time}_i + \\
&\quad\ (\alpha_{4} + \beta_{4\text{country}[i]} + \gamma_{4\text{age}[i]} + \delta_{4\text{age:country}[i]})\text{X}^{\text{CW}}_i + \\
&\quad\ (\alpha_{5} + \beta_{5\text{country}[i]} + \gamma_{5\text{age}[i]} + \delta_{5\text{age:country}[i]})\text{Sex}_i \times \text{X}^{\text{CW}}_i + \\
&\quad\ \alpha_6 \text{X}^{\text{CB}}_i,
\\
\pmb{\beta} &\sim \text{MVN}(\pmb{0}, \Sigma^\text{country}), \\
\pmb{\gamma} &\sim \text{MVN}(\pmb{0}, \Sigma^\text{age}), \\
\pmb{\delta} &\sim \text{MVN}(\pmb{0}, \Sigma^\text{age:country}),
\end{align*}

where $i$ indexes rows in the data and $\text{se}_i$ are the known standard errors. $\text{X}^{\text{CW}}_i$ and $\text{X}^{\text{CB}}_i$ are the within- and between-country centred internet technology adoption predictor (per capita internet users or mobile broadband subscriptions, in separate models), respectively. We used default noninformative priors on all parameters except the following population-level coefficients, that were modelled as

\begin{align*}
\alpha_{1,3-6} &\sim \text{Normal}(0, 1), \\
\alpha_{2} &\sim \text{Normal}(0, 3), \\
\sigma_{\gamma_4} &\sim \text{Normal}^+(0, 0.5).
\end{align*}

We used these priors to help model convergence, and on the assumption that differences in well-being between males and females are likely greater than differences associated with years, one-percent increases in technology adoption predictors, or the interaction terms.

Our third research question asked whether any trends in well-being or its associations with internet or mobile broadband adoption were specific to adolescents or other demographic groups. To answer, the model allowed all parameters to vary randomly across the age groups and the age by country interaction. Therefore, each age group, on average and within each country, received their own partially pooled estimates [@gelmanDataAnalysisUsing2007]. This bayesian approach to estimating age-specific associations is beneficial, given that the large number of age groups, especially within countries, would otherwise present difficulties with uncertain estimates and multiple comparisons. Therefore, we could compare younger age groups to older age groups, and investigate other contrasts, with confidence and without additional post-hoc adjustment procedures [@gelmanWhyWeUsually2012].

We conducted all data analyses with R [@rcoreteamLanguageEnvironmentStatistical2021] and estimated the models using Stan's Hamiltonian Monte Carlo sampling via the brms R package [@burknerBrmsPackageBayesian2017; @standevelopmentteamStanModelingLanguage2021]. We ran 4 HMC chains with 3,000 iterations and first 1,500 as warmup for 6,000 total iterations. We ensured that all population-level parameters' had converged using numeric diagnostics (all Rhats were within 0.01 of 1) and visual inspection of parameters' scatterplot matrices.

## Results

```{r data-describe}
gwp %>% 
  group_by(outcome) %>% 
  summarise(n=sum(n)) %>% 
  summarise(N=comma(median(n))) %>% 
  pull(N)

bind_rows(
  "GWP" = gwp,
  "GBD" = gbd,
  .id = "Dataset"
) %>% 
  mutate(Dataset = fct_inorder(Dataset)) %>% 
  count(Dataset, outcome) %>% 
  bind_rows(
    itu %>% 
      select(internet, mobile) %>% 
      pivot_longer(everything(), names_to = "outcome") %>% 
      drop_na(value) %>% 
      count(outcome) %>% 
      mutate(Dataset = "ITU")
  )

# Individuals using the internet per 100 inhabitants | World
if (!file.exists(
  "data-raw/ITU/ITU_regional_global_Key_ICT_indicator_aggregates_rev1_Jan_2022.xlsx")) {
  download.file("https://www.itu.int/en/ITU-D/Statistics/Documents/facts/ITU_regional_global_Key_ICT_indicator_aggregates_rev1_Jan_2022.xlsx", "data-raw/ITU/ITU_regional_global_Key_ICT_indicator_aggregates_rev1_Jan_2022.xlsx")
}
read_excel("data-raw/ITU/ITU_regional_global_Key_ICT_indicator_aggregates_rev1_Jan_2022.xlsx", range = "T85:AI86")
```

```{r models-load}
# Create a table of outcomes in appropriate order, so can map functions over this
oo <- c(
  "Life_satisfaction", "Negative_experiences", "Positive_experiences",
  "Anxiety", "Depression", "Selfharm"
)

fits <- tibble(
  path = list.files("models", pattern = "brm-", full.names = TRUE),
  outcome = basename(path)
) %>%
  separate(outcome, c("discard", "outcome", "model"), sep = "-") %>% 
  mutate(model = str_extract(model, "[0-9]")) %>% 
  filter(outcome %in% oo) %>% 
  mutate(fit = map(path, readRDS)) %>% 
  select(-c(path, discard)) %>% 
  mutate(outcome = factor(outcome, levels = oo))
fits_wb <- fits %>% 
  filter(outcome %in% oo[1:3])
fits_mh <- fits %>% 
  filter(outcome %in% oo[4:6])
rm(fits)
```

```{r gwp-model-diagnostics}
if (length(list.files("diagnostics/wb")) == 0) {
  rhats <- fits_wb %>% 
    mutate(
      out = map(
        fit, 
        ~describe_posterior(
          .x, 
          ci_method = "eti", 
          test = NULL, 
          effects = "all",
          parameters = "^cor_|^sd_|^b_"
        ) %>% 
          as_tibble() %>% 
          select(Parameter, Rhat, ESS)
      )
    ) %>% 
    select(-fit) %>% 
    unnest(out)

  # Show a table of bad rhats
  rhats %>%
    filter(!near(Rhat, 1, tol = .01))

  # Draw a scatterplot of parameters with 9 worst rhats
  # Create a tabset of this in HTML output
  rhats_sum <- rhats %>% 
    group_by(outcome, model) %>% 
    arrange(outcome, model, desc(abs(Rhat))) %>% 
    slice(1:9) %>% 
    summarise(v = list(str_c(Parameter))) %>% 
    ungroup() %>% 
    right_join(fits_wb)

  dir.create("diagnostics/wb", FALSE, recursive = TRUE)
  for (i in 1:nrow(rhats_sum)) {
    x <- slice(rhats_sum, i)
    path <- str_glue("diagnostics/wb/{x$outcome}-{x$model}.png")
    png(path, width = 1200, height = 1200, res = 120)
    pairs(x$fit[[1]]$fit, pars = x$v[[1]])
    dev.off()

  }
}
```

```{r gwp-model-process}
# Extract the relevant parameters' summaries from the models
tab_wb <- fits_wb %>% 
  mutate(
    out = map(
      fit,
      ~describe_posterior(
        .x, 
        effects = "all", 
        centrality = "mean", 
        ci = 0.95, 
        ci_method = "ETI", 
        test = "pd",
        diagnostic = NULL
      )
    )
  ) %>% 
  select(-fit) %>% 
  unnest(out) %>% 
  select(-CI)
tab_wb <- tab_wb %>% 
  # Clean parameter names
  separate(
    Parameter, 
    into = c("factor", "country", "parameter"), 
    sep = "\\[|\\]|,"
  ) %>%
  mutate(
    parameter = if_else(
      Effects == "fixed",
      str_remove(factor, "^b_"),
      parameter
    ),
    # Clean predictor name
    Predictor = case_when(
      parameter == "year" & model == "1" ~ "Time",
      parameter == "i_cw" ~ "Internet",
      parameter == "m_cw" ~ "Mobile"
    ) %>% factor(., levels = c("Time", "Internet", "Mobile"))
  ) %>% 
  mutate(
    cred = pd > .975
  )

```

Figure \@ref(fig:fig-gwp-data)A describes the levels of internet and mobile broadband adoption, and the three well-being variables over time averaged over sex and age. Figure \@ref(fig:fig-gwp-data)A shows the universal global penetration of internet technologies in the past two decades, and suggests that in contrast, changes in well-being are likely to be small.

```{r tab-gwp-estimates}
avg_wb <- tab_wb %>% 
  filter(Effects == "fixed") %>%
  mutate(
    Predictor = case_when(
      factor == "b_year" & model == "1" ~ "Time Within country",
      factor == "b_i_cw" ~ "Internet Within country",
      factor == "b_m_cw" ~ "Mobile Within country",
      factor == "b_i_cb" ~ "Internet Between countries",
      factor == "b_m_cb" ~ "Mobile Between countries"
    )
  ) %>% 
  drop_na(Predictor) %>% 
  # Create a result string for table
  mutate(
    Estimate = str_glue(
      "{number(Mean, .01)} [{number(CI_low, .01)}, {number(CI_high, .01)}]"
    ),
    pd = percent2(pd)
  ) %>% 
  select(
    Predictor, 
    Outcome = outcome, 
    Estimate,
    pd
  ) %>% 
  separate(Predictor, c("Predictor", "x"), extra = "merge") %>% 
  mutate(
    Predictor = factor(
      Predictor, 
      levels = c("Time", "Internet", "Mobile")
    )
  ) %>% 
  arrange(Predictor, Outcome)

avg_wb %>% 
  mutate(
    Estimate = str_glue("{Estimate} ({pd})"),
  ) %>% 
  select(-pd) %>% 
  pivot_wider(names_from = x, values_from = Estimate) %>% 
  group_by(Predictor) %>% 
  mutate(i = 1:n()) %>% 
  ungroup() %>% 
  mutate(
    Predictor = as.character(Predictor),
    Predictor = if_else(i==1, Predictor, ''),
    Outcome = str_replace(Outcome, "_", " ")
  ) %>% 
  select(-i) %>% 
  replace_na(list(`Between countries` = '')) %>% 
  mutate(
    across(
      3:4, 
      ~if_else(
        str_starts(., "-"), 
        ., 
        paste0("\\phantom{-}", .)
      )
    )
  ) %>% 
  apa_table(
    caption = "Well-being: Population-level estimates",
    span_text_columns = TRUE,
    escape = FALSE,
    note = "Estimates are percentage changes in outcomes associated with a one unit increase in the predictor (Time: one year; Internet and Mobile: one percent). LS: Life satisfaction, NE: Negative experiences, PE: Positive experiences. Numbers indicate posterior means, [95\\%CIs], and (posterior probabilities of direction)."
  )
```

We found that, for the average country, life satisfaction had remained relatively stable, whereas both negative and positive experiences had increased (Figure \@ref(fig:fig-gwp-data)B). Interestingly, the increase in negative experiences was nearly sixfold to that in positive experiences. Although we found a sizeable average increase in negative experiences, these results were mixed and so did not offer convincing evidence for consistent global changes in well-being during this period of global internet technology adoption, as would be expected if the latter had broad negative consequences.

We then turned to our main question; understanding the extent to which internet technology adoption predicted changes in well-being. We expanded the model of temporal changes to also predict well-being from within- and between-country centred per capita internet users and mobile broadband subscriptions (Figure \@ref(fig:fig-gwp-data)B, Within-country associations). The within-country associations are our key estimates of interest and describe the extent to which a one-percent increase in a country's per capita internet users (or mobile broadband subscriptions) predict that country's well-being and mental health, adjusting for linear temporal trends.

```{r fig-gwp-data}
#| include: true
#| fig.height: !expr ifelse(knitr::is_latex_output(), 7.6, 5)
#| fig.cap: "\\textbf{A.} Time courses of per capita internet users (row one), mobile broadband subscriptions (row two), and three psychological well-being outcomes (rows three to five). Lines indicate countries' yearly means of the respective variable, aggregated across sex and age. \\textbf{B.} Posterior probability distributions of population-level associations (green: between countries, red: within countries) between each of three well-being outcomes and time, per capita internet users, and per capita mobile broadband subscriptions. Estimates are percentage changes in outcomes associated with a one year or percent increase in the predictor. Points and lines indicate posterior means and 95\\%CIs, filled points indicate that zero is not among the 95\\% most credible parameter values."

p1 <- itu %>%
  select(
    region, country, year,
    Internet = internet,
    Mobile = mobile,
  ) %>%
  # Truncate mobile percentages for figure
  mutate(Mobile = if_else(Mobile > 150, NaN, Mobile)) %>%
  pivot_longer(
    c(Internet, Mobile),
    names_to = "outcome", values_to = "val"
  ) %>% 
  bind_rows(gwp) %>%
  mutate(outcome = fct_inorder(str_replace(outcome, "_", "_"))) %>%
  group_by(region, country, year, outcome) %>%
  summarise(mean = mean(val, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(outcome = factor(outcome, levels = c("Internet", "Mobile", oo))) %>% 
  mutate(year = make_date(year)) %>%
  ggplot(aes(year, mean)) +
  scale_y_continuous(
    "Value (%)",
    # Display percentages
    breaks = pretty_breaks(),
    expand = expansion(.05)
  ) +
  scale_x_date(
    "Year",
    breaks = make_date(c(2000, 2005, 2010, 2015, 2020)),
    date_labels = "'%y",
    expand = expansion(.02)
  ) +
  geom_line(
    aes(
      group = interaction(country, outcome)
    ),
    size = .05, col = "dodgerblue3"
  ) +
  facet_grid(
    outcome ~ region,
    scales = "free_y",
    labeller = as_labeller(function(x) str_replace(x, "_", "\n"))
  ) +
  theme(
    legend.position = "none"
  )

x <- fits_wb %>% 
  mutate(out = map(fit, ~spread_draws(.x, `^b_.*`, regex = TRUE))) %>% 
  select(-fit) %>% 
  unnest(out) %>% 
  select(outcome, model, .draw, b_year, matches("^b_[i|m]_c[w|b]$")) %>% 
  pivot_longer(starts_with("b_"), names_to = "parameter") %>% 
  mutate(
    Predictor = case_when(
      parameter == "b_year" & model == "1" ~ "Time Within country",
      parameter == "b_i_cw" ~ "Internet Within country",
      parameter == "b_m_cw" ~ "Mobile Within country",
      parameter == "b_i_cb" ~ "Internet Between countries",
      parameter == "b_m_cb" ~ "Mobile Between countries"
    )
  ) %>% 
  drop_na(Predictor, value) %>% 
  separate(Predictor, c("Predictor", "x"), extra = "merge") %>% 
  mutate(
    Predictor = factor(
      Predictor, 
      levels = c("Time", "Internet", "Mobile")
    )
  )
p2 <- x %>%   
  ggplot(aes(value, fct_rev(Predictor), fill = x)) +
  scale_color_brewer(
    "Association", palette = "Set2", aesthetics = c("fill", "color")
  ) +
  scale_shape_manual(values = c(21, 19)) +
  scale_x_continuous(
    "Estimated linear association",
    breaks = extended_breaks()
  ) +
  stat_halfeye(
    aes(shape = stat(sign(xmax) == sign(xmin))),
    position = position_dodge(.15), 
    normalize = "xy", 
    height = .8,
    point_interval = "mean_qi",
    fatten_point = 3,
    point_fill = "white",
    size = .25,
    slab_alpha = .75,
    adjust = 1.5,
    .width = .95
    ) +
  facet_wrap(
    "outcome", 
    labeller = labeller(
      .cols = as_labeller(~str_replace(., "_", " "))
    ),
    scales = "free_x"
  ) +
  theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    panel.grid.major.x = element_line(
      color = "black", linetype = "dotted", size = .1
    )
  )

(p1 / p2) + 
  plot_layout(heights = c(8, 2)) +
  plot_annotation(tag_levels = "A")
```

```{r}
fits_wb %>% 
  filter(model != 1) %>% 
  mutate(
    sigma = map_dbl(
      fit,
      ~mean(as.data.frame(.x, variable = "sigma")$sigma)
    )
  )
itu %>% 
  pivot_longer(internet:mobile) %>% 
  group_by(country, name) %>% 
  mutate(increase = value - lag(value)) %>% 
  group_by(name) %>% 
  summarise(mean_yearly_increase = mean(increase, na.rm = TRUE))
```

Internet adoption did not predict any outcome at the 95% credibility level, but a 1% increase in per capita internet users predicted a `r avg_wb[6, 4]` increase in negative experiences for the average country with `r avg_wb[6, 5]` posterior probability of direction. Mobile broadband, on the other hand, positively predicted life satisfaction (`r avg_wb[10, 4]`), with greater evidence for direction (`r avg_wb[10, 5]`). In light of substantial residual standard deviations ranging from about 6 to 8 between predictors and outcomes, and average yearly increases in internet technology adoption (internet = 2.7%, mobile = 5.8%) the observed associations were very small.

```{r fig-gwp-country}
#| cache: false
figure_country <- function(data) {
  data %>% 
    # Show countries and main relations
    filter(factor == "r_country") %>% 
    drop_na(Predictor) %>% 
    # Take out countries with only time association from this figure
    add_count(outcome, country) %>% 
    filter(n > 1) %>% 
    # Arrange per panel
    group_by(Predictor, outcome) %>% 
    arrange(Predictor, outcome, Mean) %>% 
    mutate(y = 1:n()) %>% 
    ggplot(aes(Mean, y, col = cred)) +
    scale_color_manual(values = c("dodgerblue1", "dodgerblue3")) +
    scale_y_discrete(
      "Country",
      # Fix country names
      labels = ~str_replace_all(., "\\.", " ") %>% 
        str_replace_all(., "’", "'"),
      expand = expansion(.015)
    ) +
    scale_x_continuous(
      "Estimated linear association",
      breaks = extended_breaks(6), 
      expand = expansion(c(0.01))
    ) +
    scale_shape_manual(values = c(21, 19)) +
    scale_linetype_manual(values = c(1, 1, 1)) +
    geom_linerangeh(
      size = .1,
      aes(xmin = CI_low, xmax = CI_high, linetype = cred)
    ) +
    geom_point(
      aes(shape = cred), 
      size = .55, fill = "white", stroke = .2
    ) +
    guides(
      linetype = "none", 
      shape = "none",
      color = guide_legend(override.aes = list(size = .5, stroke = 1))
    ) +
    facet_grid2(
      Predictor~outcome,
      scales = "free",
      independent = "all",
      labeller = labeller(
        .cols = as_labeller(~str_replace(., "_", " "))
      )
    ) +
    theme(
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      panel.grid.major.x = element_line(
        color = "black", linetype = "dotted", size = .1
      ),
      legend.position = "none"
    )
}
p_country_wb <- figure_country(tab_wb)
```

```{r gwp-demographics}
wb_demo <- fits_wb %>% 
  mutate(
    x = factor(model, labels = c("year", "i_cw", "m_cw")),
  ) %>% 
  mutate(
    out = map2(
      fit, x, 
      ~bind_rows(
        world = hypothesis(
          .x, 
          str_glue("{.y} + {.y}:sex1 * {c(-1, 1)} = 0")
        )$hypothesis,
        age = hypothesis(
          .x, 
          str_glue("{.y} + {.y}:sex1 * {c(-1, 1)} = 0"),
          scope = "coef", group = "age"
        )$hypothesis
      )
    )
  ) %>% 
  select(-fit) %>% 
  unnest(out) %>% 
  mutate(Sex = if_else(str_detect(Hypothesis, "-1"), "Male", "Female")) %>% 
  mutate(Group = if_else(is.na(Group), Sex, Group)) %>% 
  mutate(
    predictor = factor(
      model, 
      levels = 1:3, 
      labels = c("Time", "Internet", "Mobile")
    )
  ) %>% 
  mutate(cred = sign(CI.Lower) == sign(CI.Upper)) %>% 
  mutate(
    Group = factor(Group, levels = c("Female", "Male", levels(dat$age)))
  )
```

```{r fig-gwp-demographics}
#| cache: false
figure_demographics <- function(data) {
  data %>% 
    ggplot(aes(Estimate, fct_rev(Group), col = Sex, shape = cred)) +
    scale_color_brewer(palette = "Set2") +
    scale_shape_manual(values = c(21, 19)) +
    # Ensure zero shows
    geom_blank(aes(x = 0)) +
    scale_x_continuous(
      "Estimated linear association",
      breaks = extended_breaks(5),
      expand = expansion(.15)
    ) +
    geom_linerangeh(
      aes(xmin = CI.Lower, xmax = CI.Upper),
      size = .3,
      position = position_dodge2v(
        height = .8,
        padding = 1,
        reverse = TRUE
      )
    ) +
    geom_point(
      size = 0.9,
      fill = "white",
      position = position_dodge2v(
        height = .8,
        padding = 1,
        reverse = TRUE
      )
    ) +
    guides(shape = "none") +
    facet_grid2(
      rows = vars(predictor),
      cols = vars(outcome),
      scales = "free_x", independent = "x",
      labeller = as_labeller(~str_replace(., "_", " "))
    ) +
    theme(
      axis.title.y = element_blank(),
      legend.position = "none",
      panel.grid.major.x = element_line(
        color = "black", linetype = "dotted", size = .1
      )
    )
}

p_demo_wb <- figure_demographics(wb_demo)
```

Our model also measured the extent to which internet technology adoption predicted well-being between countries. These between-country associations indicated, with high confidence, that countries with greater average levels of internet and mobile broadband adoption tended to report greater average levels of life satisfaction and positive experiences, and lower average levels of negative experiences (Figure \@ref(fig:fig-gwp-data)B, Between-country associations). However, this pattern is likely to reflect multiple other causes impinging on both well-being and internet technology adoption.

These trends and associations are meaningful summaries of the average country among the `r length(unique(gwp$country))` examined, but do not represent specific countries which might have unique patterns of associations. Figure \@ref(fig:fig-gwp-parameters)A describes the country-level trends over time and associations linking per capita internet users and mobile broadband subscriptions to each of the three outcomes. There appeared to be no universal temporal trends that apply to all countries, whose estimates varied from significant decreases to substantial increases in well-being. Moreover, the associations between internet technology adoption and well-being varied significantly across countries, with some reporting substantial positive associations, and others large negative associations.

```{r fig-gwp-parameters}
#| cache: false
#| include: true
#| fig.height: !expr ifelse(knitr::is_latex_output(), 8.5, 6.8)
#| fig.env: "figure*"
#| fig.cap: "\\textbf{A.} Country-specific yearly changes in three well-being outcomes (Time), and their associations with per capita internet users (Internet) and mobile broadband subscriptions (Mobile). Points and lines indicate individual countries' posterior means and 95\\% CIs. Countries are sorted from the most negative to most positive estimate. Filled points indicate that the parameter's 95\\% credibility interval excludes zero (in both panels). We include a table of all countries' estimates in the online materials. \\textbf{B.} Age- and sex-specific (green: female, red: male) changes in three well-being outcomes (top row), and their associations with per capita internet users (middle row) and mobile broadband subscriptions (bottom row)."
((p_country_wb /
  p_demo_wb + theme(axis.text.y = element_text(size = rel(0.6)))) &
  theme(text = element_text(size = 8))) +
  plot_layout(heights = c(5.5, 4.5)) +
  plot_annotation(tag_levels = "A")
```

```{r tab-sds}
#| include: true
#| eval: false
tab_wb %>% 
  filter(str_starts(factor, "sd_")) %>% 
  separate(factor, c("factor", "term"), sep = "__") %>% 
  filter(factor %in% c("sd_age", "sd_country")) %>% 
  filter(
    model == "1" & term == "year" | 
      model != "1" & str_ends(term, "_cw")
  ) %>% 
  mutate(
    # years, percentages
    across(c(Mean:CI_high), ~if_else(model=="1", ., ./100)),
    across(c(Mean:CI_high), ~number(., .001)),
    res = as.character(str_glue("{Mean} [{CI_low}, {CI_high}]")),
    Outcome = str_replace(outcome, "_", " "),
    Predictor = factor(model, labels = c("Time", "Internet", "Mobile")),
    variable = case_when(
      factor == "sd_age" ~ "SD (age)",
      factor == "sd_country" ~ "SD (country)"
    )
  ) %>% 
  select(
    Predictor, Outcome, variable, res
  ) %>% 
  pivot_wider(names_from = variable, values_from = res) %>% 
  mutate(Outcome = fct_inorder(Outcome)) %>% 
  select(Predictor, Outcome, `SD (country)`, everything()) %>% 
  arrange(Predictor, Outcome) %>% 
  mutate(
    Predictor = as.character(Predictor),
    Predictor = if_else(Outcome == "Life satisfaction", Predictor, "")
  ) %>%
  apa_table(
    font_size = "scriptsize",
    placement = "htb",
    span_text_columns = FALSE,
    note = "Numbers indicate the standard deviations [95%CI] of the distributions of associations in the population of all countries."
  )
```

```{r tab-gwp-props}
wb_props <- tab_wb %>% 
  filter(factor == "r_country") %>% 
  drop_na(Predictor) %>% 
  mutate(
    Outcome = str_replace(outcome, "_", " "),
    # Predictor = as.character(Predictor),
    # Predictor = if_else(Outcome == "Life satisfaction", Predictor, "")
  )
wb_props2 <- wb_props %>% 
  mutate(
    x = case_when(
      cred & sign(Mean)==-1 ~ "Negative",
      cred & sign(Mean)==1 ~ "Positive",
      TRUE ~ "Null"
    )
  ) %>% 
  count(Outcome, Predictor, x) %>% 
  arrange(Predictor, Outcome) %>% 
  group_by(Outcome, Predictor) %>% 
  mutate(n = str_glue("{n} ({percent(n/sum(n), 1)})")) %>%  
  pivot_wider(names_from = x, values_from = n) %>% 
  mutate(across(c(Negative, Null, Positive), ~replace_na(., "0%"))) %>% 
  select(Predictor, Outcome, Negative, Null, Positive)
wb_props2 %>%   
  group_by(Predictor) %>% 
  mutate(i = 1:n()) %>% 
  ungroup() %>% 
  mutate(
    Predictor = as.character(Predictor),
    Predictor = if_else(i==1, Predictor, '')
  ) %>% 
  select(-i) %>% 
  apa_table(
    placement = "htb",
    span_text_columns = FALSE,
    caption = "Well-being: Country-specific test summary",
    note = "Numbers indicate counts and (percentages) of countries with different sign test results. Percentages may not add up to 100% because of rounding."
  )
```

To summarize these country-specific associations, we calculated proportions of countries with credibly positive and negative associations. Above we identified that internet adoption predicted small increases in negative experiences for the average country with `r avg_wb$pd[6]` confidence. Table \@ref(tab:tab-gwp-props) shows that this association was credibly positive (negative) for `r str_sub(wb_props2[5,5], 1, 2)` (`r str_sub(wb_props2[5,3], 1, 2)`) countries. `r str_sub(wb_props2[4,5], 1, 2)` (`r str_sub(wb_props2[4,3], 1, 2)`) countries had positive (negative) associations between life satisfaction and per capita internet users. The associations linking positive experiences to per capita internet users were more often inconclusive. We observed similar numbers regarding associations linking well-being to per capita mobile broadband subscriptions: `r str_sub(wb_props2[7,5], 1, 2)` (`r str_sub(wb_props2[7,3], 1, 2)`) countries had a positive (negative) association between mobile broadband adoption and life satisfaction, whereas the numbers were only `r str_sub(wb_props2[8,5], 1, 2)` (`r str_sub(wb_props2[8,3], 1, 2)`) and `r str_sub(wb_props2[9,5], 1, 2)` (`r str_sub(wb_props2[9,3], 1, 1)`) for negative and positive experiences, respectively. This heterogeneity and lack of consistent associations across countries should qualify any inferences concerning global associations and is evidence against the idea that the adoption of internet or mobile broadband has had uniform global negative effects on well-being.

These global and country-specific analyses are informative but shed no light on how internet and mobile broadband adoption might be differently associated with well-being across demographic groups. So we then examined variation across age and sex (for the average country) in the relations linking internet and mobile broadband adoption to psychological well-being. Age- and sex-specific estimates, for the average country, are shown in Figure \@ref(fig:fig-gwp-parameters)B.

```{r}
tmp <- tab_wb %>% 
  filter(
    Effects == "fixed", 
    str_detect(parameter, ":sex1"), 
    !(parameter=="year:sex1" & model != 1)
  ) %>% 
  mutate(
    x = str_glue(
      "{number(Mean, .01)} [{number(CI_low, .01)}, {number(CI_high, .01)}]"
      ), 
    .keep = "unused"
  ) %>% 
  select(outcome, parameter, x, pd)
tmp
```

We found that, for the average country, life satisfaction had increased more for females (`r tmp$x[1]`), and that both per capita internet users and mobile broadband subscriptions predicted slightly greater increases in life satisfaction for females than males (Internet: `r tmp$x[2]`; Mobile: `r tmp$x[3]`). Negative experiences had increased the least for youngest age group in our data, but slightly more overall for females than males (`r tmp$x[4]`). Similarly, positive experiences had increased most for the youngest and oldest age groups, and somewhat more overall for females (`r tmp$x[7]`). Overall, these demography-specific trends and associations indicated no clear patterns specific to a particular demographic group. Neither did they support the commonly offered narrative that young individuals, particularly young women, have experienced disproportionately large decrements in well-being in association with the adoption of internet technologies.

## Discussion

We examined the extent to which three indicators of psychological well-being had changed over time and in association with per capita internet users and mobile broadband subscriptions. Across 168 countries and 16 years, we found mixed support for changes in well-being over time and in association with internet technology adoption. Whereas life satisfaction, on average, had not changed over time or as a function of either per capita internet users, it was positively predicted by per capita mobile broadband subscriptions. We found that both negative and positive experiences had increased, but that their associations with internet technology adoption were not credibly different from zero at the 95% confidence level. Furthermore, there were large variations among these patterns of results between the countries studied, suggesting no consistent connection between internet technology adoption and country-level well-being. Finally, while some sex- and age-based differences emerged, none supported the idea that young individuals, or young women particularly, were the most at-risk group. Nevertheless, this analysis was necessarily constrained to a limited range of available outcomes reflecting subjective well-being [e.g. @jebbSubjectiveWellBeingWorld2020]. In Study 2, we extended our investigation to mental health outcomes.

# Study 2: Mental health

In the second study we extended our investigation to focus on mental health, using meta-analytic rates of anxiety, depression, and self-harm from 2000 to 2019 among 202 countries, and their associations with internet technology adoption, in place of self-reported measures of psychological well-being.

## Methods

### Mental health outcomes

We studied meta-analytic prevalence rates (per 100,000 individuals) of anxiety disorders (ICD10 F40-F44.9, F93-F93.2), depressive disorders (ICD10 F32-F33.9, F34.1), and self-harm (ICD10 X60-X64.9, X66-X84.9, Y87.0) in 204 countries from 2000 to 2019 as estimated by the Institute for Health Metrics and Evaluation's (IHME) Global Burden of Disease 2019 (GBD) study [@jamesGlobalRegionalNational2018; @vosGlobalBurden3692020]. The GBD collates heterogeneous data from all WHO member states' censuses, household surveys, civil registration and vital statistics, disease registries, health service use statistics, disease notifications, and other sources. It then aggregates data from these sources with bayesian meta-regression to produce country-specific yearly prevalence estimates.

The GBD 2019 prevalence rate estimates are based on 19,773 data sources with varying coverage for individual countries; for details of the GBD 2019 methodology, see [@vosGlobalBurden3692020] and especially Appendix 1 therein. The prevalence rates are estimated for females and males in 5-year age groups, and are provided as the IHME meta-regression model's predicted rates and 95% credibility intervals; we converted the latter to approximate standard errors for our meta-analytic modelling strategy (see Study 1, Methods).

We emphasize that the GBD estimates are not observed data, and therefore are accurate only to the extent that the GBD's data collection methods and modelling strategies are valid. We have compared the GBD estimates to the CDC's estimates of self-harm in the United States [@u.s.centersfordiseasecontrolandpreventionWISQARSFatalInjury2022], and found that they are likely to deviate in systematic ways from other authoritative information sources. We nevertheless argue that because the GBD provides the most comprehensive dataset of global mental health, studying these estimates is informative, but emphasize this caveat. The sample size for our analyses (combinations of country, year, sex, and age) was 130,560.

### Data analysis

We analysed the data in the same manner as in Study 1, but did not include varying parameters over the age by country interaction because the models did not converge due to invariance in the data. We also set the observed standard errors in the outcomes to zero for models of self-harm due to convergence problems.

```{r gbd-model-diagnostics}
if (length(list.files("diagnostics/mh")) == 0) {
  rhats <- fits_mh %>% 
    mutate(
      out = map(
        fit, 
        ~describe_posterior(
          .x, 
          ci_method = "eti", 
          test = NULL, 
          effects = "all",
          parameters = "^cor_|^sd_|^b_"
        ) %>% 
          as_tibble() %>% 
          select(Parameter, Rhat, ESS)
      )
    ) %>% 
    select(-fit) %>% 
    unnest(out)

  # Show a table of bad rhats
  rhats %>%
    filter(!near(Rhat, 1, tol = .01))

  # Draw a scatterplot of parameters with 9 worst rhats
  # Create a tabset of this in HTML output
  rhats_sum <- rhats %>% 
    group_by(outcome, model) %>% 
    arrange(outcome, model, desc(abs(Rhat))) %>% 
    slice(1:9) %>% 
    summarise(v = list(str_c(Parameter))) %>% 
    ungroup() %>% 
    right_join(fits_mh)

  dir.create("diagnostics/mh", FALSE, recursive = TRUE)
  for (i in 1:nrow(rhats_sum)) {
    x <- slice(rhats_sum, i)
    path <- str_glue("diagnostics/mh/{x$outcome}-{x$model}.png")
    png(path, width = 1200, height = 1200, res = 120)
    pairs(x$fit[[1]]$fit, pars = x$v[[1]])
    dev.off()

  }
}
```

```{r gbd-model-process}
# Extract the relevant parameters' summaries from the models

tab_mh <- fits_mh %>% 
  mutate(
    out = map(
      fit,
      ~describe_posterior(
        .x, 
        effects = "all", 
        centrality = "mean", 
        ci = 0.95, 
        ci_method = "ETI", 
        test = "pd",
        diagnostic = NULL
      )
    )
  ) %>% 
  select(-fit) %>% 
  unnest(out) %>% 
  select(-CI)
tab_mh <- tab_mh %>% 
  # Clean parameter names
  separate(
    Parameter, 
    into = c("factor", "country", "parameter"), 
    sep = "\\[|\\]|,"
  ) %>%
  mutate(
    parameter = if_else(
      Effects == "fixed",
      str_remove(factor, "^b_"),
      parameter
    ),
    # Clean predictor name
    Predictor = case_when(
      parameter == "year" & model == "1" ~ "Time",
      parameter == "i_cw" ~ "Internet",
      parameter == "m_cw" ~ "Mobile"
    ) %>% factor(., levels = c("Time", "Internet", "Mobile"))
  ) %>% 
  mutate(
    cred = pd > .975
  )

```

## Results

```{r tab-gbd-estimates}
avg_mh <- tab_mh %>% 
  filter(Effects == "fixed") %>%
  mutate(
    Predictor = case_when(
      factor == "b_year" & model == "1" ~ "Time Within country",
      factor == "b_i_cw" ~ "Internet Within country",
      factor == "b_m_cw" ~ "Mobile Within country",
      factor == "b_i_cb" ~ "Internet Between countries",
      factor == "b_m_cb" ~ "Mobile Between countries"
    ) %>% fct_inorder
  ) %>% 
  drop_na(Predictor) %>% 
  # Models were estimated with percentages, change to rates per 100k
  mutate(
    across(Mean:CI_high, ~ . * 1000)
  ) %>% 
  # Create a result string for table
  mutate(
    Estimate = str_glue(
      "{number(Mean, .01)} [{number(CI_low, .01)}, {number(CI_high, .01)}]"
    ),
    pd = percent2(pd)
  ) %>% 
  select(
    Predictor, 
    Outcome = outcome, 
    Estimate,
    pd
  ) %>% 
  arrange(Predictor, Outcome) %>% 
  separate(Predictor, c("Predictor", "x"), extra = "merge")

avg_mh %>% 
  mutate(
    Estimate = str_glue("{Estimate} ({pd})"),
  ) %>% 
  select(-pd) %>% 
  pivot_wider(names_from = x, values_from = Estimate) %>% 
  group_by(Predictor) %>% 
  mutate(i = 1:n()) %>% 
  ungroup() %>% 
  mutate(
    Predictor = as.character(Predictor),
    Predictor = if_else(i==1, Predictor, ''),
    Outcome = str_replace(Outcome, "_", " ")
  ) %>% 
  select(-i) %>% 
  mutate(Outcome = str_replace(Outcome, "Selfharm", "Self-harm")) %>% 
  replace_na(list(`Between countries` = '')) %>% 
  mutate(
    across(
      3:4, 
      ~if_else(
        str_starts(., "-"), 
        ., 
        paste0("\\phantom{-}", .)
      )
    )
  ) %>% 
  apa_table(
    # col.names = c("X", "Y", "Within country", "Between country"),
    caption = "Mental health: Population-level estimates",
    span_text_columns = TRUE,
    # font_size = "scriptsize",
    escape = FALSE,
    note = "Estimates are changes in outcome rates associated with a one unit increase in the predictor (Time: one year; Internet and Mobile: one percent). Numbers indicate posterior means, [95\\%CIs], and (posterior probabilities of direction)."
  )
```

Because Study 2 concerned GBD's meta-analytic estimates of mental health rates, we observed much less variation in the variables over time than was the case with the well-being outcomes (Figure \@ref(fig:fig-gbd-data)A). Our results showed that rates of anxiety had increased (`r avg_mh$Estimate[1]`), whereas those of depression (`r avg_mh$Estimate[2]`) and self-harm (`r avg_mh$Estimate[3]`) had decreased, and the model certainty in each of these average trends was high (>99.9%).

```{r fig-gbd-data}
#| include: true
#| fig.height: !expr ifelse(knitr::is_latex_output(), 6.6, 5)
#| fig.cap: "\\textbf{A.} Time courses of GBD's meta-analytic country-specific estimates of rates of anxiety, depression, and self-harm. Lines indicate countries' yearly means of the respective variable, aggregated across sex and age.  \\textbf{B.} Posterior probability distributions of population-level associations (green: between countries, red: within countries) between each of three mental health outcomes and time, per capita internet users, and per capita mobile broadband subscriptions. Estimates are percentage changes in outcomes associated with a one year or percent increase in the predictor. Points and lines indicate posterior means and 95\\%CIs, filled points indicate that zero is not among the 95\\% most credible parameter values."

p1 <- gbd %>% 
  mutate(outcome = fct_inorder(outcome)) %>%
  group_by(region, country, year, outcome) %>%
  summarise(mean = mean(val, na.rm = TRUE)) %>%
  ungroup() %>% 
  mutate(year = make_date(year)) %>%
  mutate(outcome = str_replace(outcome, "Selfharm", "Self-harm")) %>% 
  ggplot(aes(year, mean)) +
  scale_y_continuous(
    "Rate per 100k individuals",
    # From percentages in model to rates per 100k
    labels = ~ number(. * 1000, big.mark = ","),
    breaks = pretty_breaks(5),
    expand = expansion(.05)
  ) +
  scale_x_date(
    "Year",
    breaks = make_date(c(2000, 2005, 2010, 2015, 2020)),
    date_labels = "'%y",
    expand = expansion(.02)
  ) +
  geom_line(
    aes(
      group = interaction(country, outcome)
    ),
    size = .05, col = "dodgerblue3"
  ) +
  facet_grid(
    outcome ~ region,
    scales = "free_y",
    labeller = as_labeller(function(x) str_replace(x, "_", "\n"))
  ) +
  theme(
    legend.position = "none"
  )

x <- fits_mh %>% 
  mutate(out = map(fit, ~spread_draws(.x, `^b_.*`, regex = TRUE))) %>% 
  select(-fit) %>% 
  unnest(out) %>% 
  select(outcome, model, .draw, b_year, matches("^b_[i|m]_c[w|b]$")) %>% 
  pivot_longer(starts_with("b_"), names_to = "parameter") %>% 
  mutate(
    Predictor = case_when(
      parameter == "b_year" & model == "1" ~ "Time Within country",
      parameter == "b_i_cw" ~ "Internet Within country",
      parameter == "b_m_cw" ~ "Mobile Within country",
      parameter == "b_i_cb" ~ "Internet Between countries",
      parameter == "b_m_cb" ~ "Mobile Between countries"
    )
  ) %>% 
  drop_na(Predictor, value) %>% 
  separate(Predictor, c("Predictor", "x"), extra = "merge") %>% 
  mutate(
    Predictor = factor(
      Predictor, 
      levels = c("Time", "Internet", "Mobile")
    )
  )
p2 <- x %>%   
  ggplot(aes(value, fct_rev(Predictor), fill = x)) +
  scale_color_brewer(
    "Association", palette = "Set2", aesthetics = c("fill", "color")
  ) +
  scale_shape_manual(values = c(21, 19)) +
  scale_x_continuous(
    "Estimated linear association",
    breaks = extended_breaks(),
    # From percentages in model to rates per 100k
    labels = ~ . * 1000,
  ) +
  stat_halfeye(
    aes(shape = stat(sign(xmax) == sign(xmin))),
    position = position_dodge(.15), 
    normalize = "xy", 
    height = .8,
    point_interval = "mean_qi",
    fatten_point = 3,
    point_fill = "white",
    size = .25,
    slab_alpha = .75,
    adjust = 1.5,
    .width = .95
    ) +
  facet_wrap(
    "outcome", 
    labeller = labeller(
      .cols = as_labeller(~str_replace(., "_", " "))
    ),
    scales = "free_x"
  ) +
  theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    panel.grid.major.x = element_line(
      color = "black", linetype = "dotted", size = .1
    )
  )

(p1 / p2) + 
  plot_layout(heights = c(7, 3)) +
  plot_annotation(tag_levels = "A")
```

Then, to answer our primary research question, we examined the within-country predictive associations between internet technology adoption and mental health. We observed no credible (at the 95% level) associations between internet adoption and either anxiety, depression, or self-harm (Figure \@ref(fig:fig-gbd-data)B). Similarly, per capita mobile broadband subscriptions were not credible predictors of mental health for the average country, although the model indicated `r avg_mh$pd[10]` certainty of a negative association between mobile broadband adoption and rates of anxiety.

We also examined the between-country associations linking internet technology adoption with mental health. Countries with greater levels of per capita internet users, on average, tended to be those with lower levels of depression, and higher levels of self-harm (Figure \@ref(fig:fig-gbd-data)B, Between country estimates). Similarly, per capita mobile broadband subscriptions predicted lower rates of both anxiety and depression, but higher rates of self-harm.

We then examined the heterogeneity in the associations between countries. Although there were considerable differences in changes in anxiety, depression, and self-harm between countries, the associations varied less (Figure \@ref(fig:fig-gbd-parameters)A). We found that internet adoption was positively (negatively) associated with anxiety for three (seven) countries, with depression for one (zero) countries, and self-harm for seven (ten) countries (Table \@ref(tab:tab-gbd-props)). We include all countries' estimates in the online materials.

```{r fig-gbd-country}
#| cache: false
p_country_mh <- figure_country(
  tab_mh %>% 
    mutate(outcome = str_replace(outcome, "Selfharm", "Self-harm"))
) +
  scale_x_continuous(
    "Estimated linear association",
    # From percentages in model to rates per 100k
    labels = ~ . * 1000,
    breaks = extended_breaks(6), 
    expand = expansion(c(0.01))
  )
```

```{r gbd-demographics}
mh_demo <- fits_mh %>% 
  mutate(
    x = factor(model, levels = 1:3, labels = c("year", "i_cw", "m_cw")),
  ) %>% 
  mutate(
    out = map2(
      fit, x, 
      ~bind_rows(
        world = hypothesis(
          .x, 
          str_glue("{.y} + {.y}:sex1 * {c(-1, 1)} = 0")
        )$hypothesis,
        age = hypothesis(
          .x, 
          str_glue("{.y} + {.y}:sex1 * {c(-1, 1)} = 0"),
          scope = "coef", group = "age"
        )$hypothesis
      )
    )
  ) %>% 
  select(-fit) %>% 
  unnest(out) %>% 
  mutate(Sex = if_else(str_detect(Hypothesis, "-1"), "Male", "Female")) %>% 
  mutate(Group = if_else(is.na(Group), Sex, Group)) %>% 
  mutate(
    predictor = factor(
      model, 
      levels = 1:3, 
      labels = c("Time", "Internet", "Mobile")
    )
  ) %>% 
  mutate(cred = sign(CI.Lower) == sign(CI.Upper)) %>% 
  mutate(
    Group = factor(Group, levels = c("Female", "Male", levels(dat$age)))
  ) %>% 
  mutate(outcome = str_replace(outcome, "Selfharm", "Self-harm"))
```

```{r fig-gbd-parameters}
#| include: true
#| #| cache: false
#| fig.height: !expr ifelse(knitr::is_latex_output(), 8.5, 6.8)
#| fig.env: "figure*"
#| fig.cap: "\\textbf{A.} Country-specific yearly changes in three mental health outcomes (Time), and their associations with per capita internet users (Internet) and mobile broadband subscriptions (Mobile). Points and lines indicate individual countries' posterior means and 95\\% CIs. Countries are sorted from the most negative to most positive estimate. Filled points indicate that the parameter's 95\\% credibility interval excludes zero (in both panels). We include a table of all countries' estimates in the online materials. \\textbf{B.} Age- and sex-specific (green: female, red: male) changes in three mental health outcomes (top row), and their associations with per capita internet users (middle row) and mobile broadband subscriptions (bottom row)."
p_demo_mh <- figure_demographics(mh_demo) +
  # Adjust x-axes
  scale_x_continuous(
    "Estimated linear association",
    breaks = extended_breaks(5),
    expand = expansion(.1),
    labels = ~. * 1000
  )

((p_country_mh /
  p_demo_mh + theme(axis.text.y = element_text(size = rel(0.6)))) &
  theme(text = element_text(size = 8))) +
  plot_layout(heights = c(5.5, 4.5)) +
  plot_annotation(tag_levels = "A")
```

```{r tab-gbd-props}
mh_props <- tab_mh %>% 
  filter(factor == "r_country") %>% 
  drop_na(Predictor) %>% 
  mutate(
    Outcome = str_replace(outcome, "_", " "),
    # Predictor = as.character(Predictor),
    # Predictor = if_else(Outcome == "Life satisfaction", Predictor, "")
  ) %>% 
  mutate(outcome = str_replace(Outcome, "Selfharm", "Self-harm"))
mh_props2 <- mh_props %>% 
  mutate(
    x = case_when(
      cred & sign(Mean)==-1 ~ "Negative",
      cred & sign(Mean)==1 ~ "Positive",
      TRUE ~ "Null"
    )
  ) %>% 
  count(Outcome, Predictor, x) %>% 
  arrange(Predictor, Outcome) %>% 
  group_by(Outcome, Predictor) %>% 
  mutate(n = str_glue("{n} ({percent(n/sum(n), 1)})")) %>%  
  pivot_wider(names_from = x, values_from = n) %>% 
  mutate(across(c(Negative, Null, Positive), ~replace_na(., "0%"))) %>% 
  select(Predictor, Outcome, Negative, Null, Positive)
mh_props2 %>%   
  group_by(Predictor) %>% 
  mutate(i = 1:n()) %>% 
  ungroup() %>% 
  mutate(
    Predictor = as.character(Predictor),
    Predictor = if_else(i==1, Predictor, '')
  ) %>% 
  select(-i) %>% 
  apa_table(
    placement = "htb",
    span_text_columns = FALSE,
    caption = "Mental health: Country-specific test summary",
    note = "Numbers indicate counts and (percentages) of countries with different sign test results. Percentages may not add up to 100% because of rounding."
  )
```

```{r}
tmp <- tab_mh %>% 
  filter(
    Effects == "fixed", 
    str_detect(parameter, ":sex1"), 
    !(parameter=="year:sex1" & model != 1)
  ) %>% 
  mutate(
    x = str_glue("{percent(Mean, .01)} [{number(CI_low, .01)}, {number(CI_high, .01)}]"), 
    .keep = "unused"
  ) %>% 
  select(outcome, parameter, x, pd)
tmp
```

We then turned to our third research question, and studied how the associations linking internet technology adoption and mental health varied across demographic groups. Changes in the rates of anxiety showed a complex pattern, whereby older males rates of anxiety increased more than younger males, except the very youngest age group (Figure \@ref(fig:fig-gbd-parameters)B). For females, on the other hand, rates of anxiety increased least for the oldest and youngest groups. Overall, the difference in changes in anxiety between males and females was not credibly different from zero. Depression had decreased more for females than males (`r tmp$x[4]`), and for the middle age groups vs. the youngest and oldest groups. Self-harm had decreased slightly more for females, but the difference was not credibly different from zero, and for older groups vs. younger.

Overall, the demography-based differences in the mental health-internet adoption associations were smaller than within the mental health trends (Figure \@ref(fig:fig-gbd-parameters)B). There were no sex differences in associations linking either per capita internet users or mobile broadband subscriptions to either mental health outcome that were credibly different from zero. Age-based differences in links with anxiety were small. However, depression was negatively and credibly associated with internet adoption for older age groups, but positively and credibly associated with per capita mobile broadband subscriptions for both sexes in the 10 to 14 and 15 to 19 year-old age brackets. Similarly, per capita internet users predicted self-harm negatively for older age groups in the 65 to 89 year-old brackets, but positively in the 10 to 34 year-old brackets. These results indicated some support for the idea that younger individuals' mental health may be more negatively associated with the adoption of internet technologies, although this finding was not consistent across the two different predictors or three different mental health outcomes.

## Discussion

In Study 2, we replicated the analyses regarding well-being in Study 1 using three mental health outcomes from a different dataset. We found mixed results regarding changes in the rates of mental health problems, whereby rates of anxiety had increased, but rates of depression and self-harm decreased, for the average country. Importantly, demography-based differences emerged: Younger age groups' rates of depression and self-harm were more positively predicted by both per capita internet users and mobile broadband subscriptions.

# General discussion

The idea that the rapid and global penetration of the internet, and technologies enabled by it, is affecting psychological well-being and mental health is compelling but not adequately tested. In two studies, we focused on the degree to which well-being and mental health had changed in the past two decades across most world countries, and found mixed results: Both negative and positive experiences had increased, as had meta-analytic rates of anxiety. Depression and self-harm, conversely, had decreased. These results align with other results that tend to be mixed or find few changes over time [@borAreChildAdolescent2014; @collishawAnnualResearchReview2015; @cosmaCrossNationalTimeTrends2020; @keyesRecentIncreasesDepressive2019]. In addition, although previous reports have suggested increasing mental health problems specific to young females [e.g. @borAreChildAdolescent2014], we did not find support for such changes for the average country.

We then measured the extent to which countries' per capita internet users and mobile broadband subscriptions predicted those countries' levels of well-being and mental health. We found that, overall, there were few if any meaningfully large average associations. In all cases but one---mobile broadband adoption predicted life satisfaction positively---these associations were not credibly different from zero (Figure \@ref(fig:fig-gwp-data)B). The small observed associations were also qualified by substantive variability across countries (Figure \@ref(fig:fig-gwp-parameters)A) and demographics (Figure \@ref(fig:fig-gwp-parameters)B). Furthermore, countries where internet technologies were more widely adopted overall were not those with lowest levels of well-being and mental health.

Finally, we found large age-based differences in associations linking internet technology adoption and mental health outcomes. Both depression and self-harm were more positively associated with per capita internet users and mobile broadband subscriptions in the younger age brackets---including the youngest studied here at 10 to 14 years---than they were for older age groups. Studies on the association between internet technology use and mental health across wide age ranges are scarce. Nevertheless, while we found few gender differences, our results are in line with one study reporting that cross-sectional associations between social media use and life satisfaction are more negative at specific time windows in adolescence [@orbenWindowsDevelopmentalSensitivity2022].

Our results do not provide evidence supporting the view that the internet, and technologies enabled by it such as smartphones with internet access, are actively promoting or harming either well-being or mental health globally. Our results also suggest that age- and sex-based differences in associations linking internet technology adoption and well-being are small. However, we did find evidence suggesting that internet technology adoption is more negatively associated with young individuals' mental health than older groups' (whose associations were, in some cases, positive).

It is important to note that the data and theory required to address this question at the causal level are absent. Consequently, our analyses cannot account for potential confounders in the associations linking mental health and well-being to internet technology adoption. Our descriptions, therefore, are suggestive but are not intended to provide evidence for, or against, causal relations. Although the first to study this topic within a global perspective, future studies should investigate other potential differences based on demographic, economic, and cultural factors, which we could not address here. In addition, our results regarding mental health, while comprehensive, are necessarily and considerably less certain than those on well-being, due to the lack of diverse global data and our own reliance on GBD estimates in place of objective measurements of mental health.

Research on the effects of internet technologies is stalled because the data most urgently needed are collected and held behind closed doors by technology companies and online platforms. It is crucial to study, in more detail and with more transparency from all stakeholders, data on individual adoption of and engagement with internet-based technologies. These data exist and are continuously analysed by global technology firms for marketing and product improvement, but unfortunately are not accessible for independent research. It remains a fundamental challenge to this field of inquiry to ensure that this information is accessible to independent scholars. Until these data can be transparently analysed for the public good, the potential harmful effects of the internet and other digital environments will remain unknown.

# Material availability and ethics

All the data-analytic code and a synthetic GWP dataset to facilitate the verification, extension, and critique of our procedures are available at <https://doi.org/10.5281/zenodo.7004053>. The ITU data, along with a complete list of years, countries, and sources of data are freely available at <https://www.itu.int/en/ITU-D/Statistics/Pages/stat/default.aspx>. The GWP data are available to subscribing institutions through Gallup. The GBD data are freely available at <http://ghdx.healthdata.org/gbd-results-tool>.

This study and methods therein were approved by the University of Oxford Central University Research Ethics Committee (SSH_OII_CIA_21_084).

# Author contributions

MV: Conceptualization, data curation, formal analysis, investigation, methodology, project administration, software, validation, visualization, writing - original draft preparation, writing - review & editing. AKP: Conceptualization, funding acquisition, investigation, methodology, project administration, resources, supervision, visualization, writing - original draft preparation, writing - review & editing.

# Conflicts of interest

The authors declare no competing interests. The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.

# Acknowledgements

We thank Niklas Johannes for feedback on this work. The authors would like to acknowledge the use of the University of Oxford Advanced Research Computing (ARC) facility in carrying out this work http://dx.doi.org/10.5281/zenodo.22558.

# Funding

This research was supported by the Huo Family Foundation and Economic and Social Research Council Grant ES/T008709/1.

# References

::: {#refs}
:::

\clearpage
\onecolumn

\begin{center}
\textbf{Supplementary online material to \emph{Global well-being and mental health in the internet age} (Vuorre \& Przybylski)}
\end{center}

<!-- Prepend table & figure numbers with S and reset counters -->
\setcounter{page}{1}
\setcounter{table}{0}
\setcounter{figure}{0}
\renewcommand{\thetable}{SOM\arabic{table}}
\renewcommand{\thefigure}{SOM\arabic{figure}}

## Well-being: Country-specific parameters

```{r, tab-gwp-props}
#| include: !expr knitr::is_latex_output()
```

```{r}
#| include: !expr knitr::is_latex_output()
table_of_countries <- function(data, x) {

  # Display MH as rates per 100k individuals
  m <- if_else(x == "Mental health", 1000, 1)

  data %>% 
    filter(factor == "r_country") %>% 
    drop_na(Predictor) %>% 
    mutate(
      across(Mean:CI_high, ~. * m),
      Estimate = str_glue(
        "{number(Mean, .01)} [{number(CI_low, .01)}, {number(CI_high, .01)}] ({percent2(pd)})"
      ),
      outcome = str_replace(outcome, "_", " ")
    ) %>% 
    select(
      Outcome = outcome,
      Predictor, 
      Country = country,
      Estimate
    ) %>% 
    pivot_wider(names_from = Outcome, values_from = Estimate) %>% 
    mutate(
      across(
        2:5, 
        ~if_else(
          str_starts(., "-"), 
          ., 
          paste0("\\phantom{-}", .)
        )
      ),
      across(2:5, ~ifelse(is.na(.), "", .))
    ) %>% 
    arrange(Predictor, Country) %>% 
    mutate(Country = str_replace_all(Country, "\\.(?!\\.)", " ")) %>% 
    apa_table(
      longtable = TRUE, 
      escape = FALSE,
      font_size = "scriptsize",
      caption = str_glue("{x}: Country-specific estimates"),
      note = "Estimates are changes in the outcome associated with a one unit increase in the predictor (Time: one year; Internet and Mobile: one percent). Numbers indicate posterior means, [95\\%CIs], and (posterior probabilities of direction)."
    )
}
table_of_countries(tab_wb, "Well-being" )
```

## Mental health: Country-specific parameters

```{r, tab-gbd-props}
#| include: !expr knitr::is_latex_output()
```

```{r}
#| include: !expr knitr::is_latex_output()
table_of_countries(tab_mh, "Mental health" )
```