### 2023-05-30 meta-regression-weights.qmd
Include cell Ns as weights

---
title: "Meta-regression weighting"
format: html
---

Confirming how brms syntax works re meta-regression.

```{r}
library(brms)
library(janitor)
library(posterior)
library(ggstance)
library(tidyverse)
```

Get a snippet of data

```{r}
dat <- read_rds("data-raw/Gallup/gwp-processed.rds") %>%
  clean_names() %>%
  transmute(
    country = countrynew,
    year = year_calendar,
    id = wpid,
    sex = factor(wp1219, levels = c(2, 1), labels = c("Female", "Male")),
    age = wp1220,
    # Don't know, refused, and missing values
    Life_satisfaction = if_else(between(wp16, 0, 10), wp16, NaN),
    across(wp60:wp74, ~ if_else(between(., 1, 2), ., NaN)),
    # Also reverse the weird 1: yes 2: no coding here
    across(wp60:wp74, ~ 3 - .)
  ) %>% 
  filter(country == "United States", year == 2015) %>% 
  select(sex, age, Life_satisfaction) %>% 
  filter(age <= 89) %>% 
  mutate(
    age = cut(
      age,
      breaks = seq(15, 90, by = 5),
      include.lowest = TRUE,
      right = FALSE,
      labels = paste(seq(15, 85, by = 5), "to", seq(19, 89, by = 5))
    ) %>% as.character()
  )
dat_sum <- dat %>% 
  summarise(
    n = sum(!is.na(Life_satisfaction)),
    sd = sd(Life_satisfaction, na.rm = TRUE),
    se = sd / sqrt(n),
    Life_satisfaction = mean(Life_satisfaction, na.rm = TRUE), 
    .by = c(age, sex)
    )
```

Specify models

```{r}
# Raw data
f1 <- brm(
  Life_satisfaction ~ 0 + age %in% sex,
  data = dat,
  backend = "cmdstanr",
  cores = 4
)

# With SEs
f2.1 <- brm(
  Life_satisfaction | se(se, sigma = TRUE) ~ 0 + age %in% sex,
  data = dat_sum,
  backend = "cmdstanr",
  cores = 4
)
f2.2 <- brm(
  Life_satisfaction | se(se, sigma = TRUE) + weights(n) ~ 0 + age %in% sex,
  data = dat_sum,
  backend = "cmdstanr",
  cores = 4
)

# With SDs
f3.1 <- brm(
  Life_satisfaction | se(sd, sigma = TRUE) ~ 0 + age %in% sex,
  data = dat_sum,
  backend = "cmdstanr",
  cores = 4
)
f3.2 <- brm(
  Life_satisfaction | se(sd, sigma = TRUE) + weights(n) ~ 0 + age %in% sex,
  data = dat_sum,
  backend = "cmdstanr",
  cores = 4
)
```

Compare estimates

```{r}
bind_rows(
  as_draws_df(f1) %>% 
    summarize_draws(),
  as_draws_df(f2.1) %>% 
    summarize_draws(),
  as_draws_df(f2.2) %>% 
    summarize_draws(),
  as_draws_df(f3.1) %>% 
    summarize_draws(),
  as_draws_df(f3.2) %>% 
    summarize_draws(),
  .id = "Model"
) %>% 
  mutate(
    Model = factor(Model, labels = c("1", "2.1", "2.2", "3.1", "3.2")),
    sex = factor(str_detect(variable, "sexFemale"), labels = c("Male", "Female"))
  ) %>% 
  filter(str_starts(variable, "b_")) %>% 
  ggplot(aes(mean, variable, col = Model)) +
  geom_pointrangeh(
    aes(xmin = q5, xmax = q95),
    position = position_dodgev(.75)
  ) +
  facet_wrap("sex", scales = "free_y")
```

Standardizing

```{r}
f1.2 <- brm(
  scale(Life_satisfaction) ~ 0 + age %in% sex,
  data = dat,
  backend = "cmdstanr",
  cores = 4
)
```

```{r}
bind_rows(
  as_draws_df(f1) %>% 
    summarize_draws(),
  as_draws_df(f1.2) %>% 
    summarize_draws(),
  .id = "Model"
) %>% 
  filter(str_starts(variable, "b_")) %>% 
  mutate(
    across(
      where(is.numeric), 
      ~if_else(
        Model == 1, 
        (. - mean(dat$Life_satisfaction, na.rm = TRUE)) / sd(dat$Life_satisfaction, na.rm = TRUE), 
        .
      )
    )
  ) %>% 
  ggplot(aes(mean, variable, col = Model)) +
  geom_pointrangeh(
    aes(xmin = q5, xmax = q95),
    position = position_dodgev(1)
  )
```